---
title: "PCIT-CoS-Waitlist RCT Paper 1: DPICS NTA"
author: "Nancy Briggs"
date: "05 JAN 2021"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---


<!-- ### DPICS NTA  {.tabset .tabset-fade .tabset-pills} -->
### DPICS NTA 

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(lme4)
library(tidyverse)
library(tidyselect)
library(SASmarkdown)
library(nlme)
library(emmeans)
library(multcomp)
library(kableExtra)
library(ggplot2)
library(lmerTest)
library(DHARMa)

options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)

```


```{r, echo=FALSE}
# Data 
# Import data
# recoding for CBCL and totals done in SPSS before bringing data in to R

data <- read.csv("../data_prep.csv", fileEncoding="UTF-8-BOM")
names(data) <- tolower(names(data))

# keep relevant demographic and outcome variables

data <- data %>%
  dplyr::select(id:treatment_time_mins, dobchild, sexchild, numbhousehold, incfamily,
                #DPICS Labelled Praise
                parent_lp_perc_total_time_1, parent_lp_perc_total_time_2, parent_lp_perc_total_time_3,
                #DPICS Negative Talk
                parent_nta_perc_total_time_1, parent_nta_perc_total_time_2, parent_nta_perc_total_time_3,
                #BCAP
                prebcap_risk, post_bcap_risk, fu_bcap_risk,
                #BCAP Lie scale
                prebcap_lie, post_bcap_lie, fu_bcap_lie,
                
                # Child Ext Behav
                cbclext_pre_total,cbclext_post_total,cbclext_fu_total,

                # CBCL Internalising Behav
                cbclint_pre_total,cbclint_post_total,cbclint_fu_total,
                
                #Child Compliance
                total_co_perc_time_1,total_co_perc_time_2,total_co_perc_time_3
                
                # others, but not essential for paper
                #3. Child Temperament (ECBQ)
                #4. Child social-emotional functioning - #1: Brief Infant Toddler Social Emotional Assessment (BITSEA)
                #5. Child social-emotional functioning - #2: Child Devereux Early Childhood Assessment (DECA)
                
                )


### import total time
time <- read.csv("../data_prep_DPiCStime.csv", fileEncoding="UTF-8-BOM")
names(time) <- tolower(names(time))
time <- time %>%
  dplyr::select(id, parent_total_verbalisations_time_1, parent_total_lp_time_1, 
                  parent_total_nta_time_1, total_eff_commands_time_1, 
                  child_co_total_time_1, child_nc_total_time_1, 
                  total_co_perc_time_1, total_nc_perc_time_1, 
                  parent_total_verbalisations_time_2, parent_total_lp_time_2, 
                  parent_total_nta_time_2, total_eff_commands_time_2, 
                  child_co_total_time_2, child_nc_total_time_2, 
                  total_co_perc_time_2, total_nc_perc_time_2, 
                  parent_total_verbalisations_time_3, parent_total_lp_time_3, 
                  parent_total_nta_time_3, total_eff_commands_time_3, 
                  child_co_total_time_3, child_nc_total_time_3, 
                  total_co_perc_time_3, total_nc_perc_time_3	)

# Join
data <- left_join(data, time, by="id")


```




#### Descriptives

```{r, echo=FALSE}
# parent_nta_perc_total_time_1, parent_nta_perc_total_time_2, parent_nta_perc_total_time_3,


desctable <- data %>%
  group_by(group) %>%
  mutate(score1 = case_when( 
    parent_nta_perc_total_time_1 == -9 ~ NA_real_,                                #assign missing values to zeros
    TRUE ~ parent_nta_perc_total_time_1)) %>%
  mutate(score2 = case_when( 
    parent_nta_perc_total_time_2 == -9 ~ NA_real_,                                #assign missing values to zeros
    TRUE ~ parent_nta_perc_total_time_2)) %>%
  mutate(score3 = case_when( 
    parent_nta_perc_total_time_3 == -9 ~ NA_real_,                                #assign missing values to zeros
    TRUE ~ parent_nta_perc_total_time_3)) %>%
  summarise(n1 = sum(!is.na(score1)) , 
            mean1=mean(parent_nta_perc_total_time_1, na.rm = T ),
            sd1=sd(parent_nta_perc_total_time_1, na.rm = T ),
            median1=median(parent_nta_perc_total_time_1, na.rm = T ),
            min1=min(parent_nta_perc_total_time_1, na.rm = T ),
            max1=max(parent_nta_perc_total_time_1, na.rm = T ),
            
            n2 = sum(!is.na(score2)) , 
            mean2=mean(parent_nta_perc_total_time_2, na.rm = T ),
            sd2=sd(parent_nta_perc_total_time_2, na.rm = T ),
            median2=median(parent_nta_perc_total_time_2, na.rm = T ),
            min2=min(parent_nta_perc_total_time_2, na.rm = T ),
            max2=max(parent_nta_perc_total_time_2, na.rm = T ),
            
            n3 = sum(!is.na(score3)) , 
            mean3=mean(parent_nta_perc_total_time_3, na.rm = T ),
            sd3=sd(parent_nta_perc_total_time_3, na.rm = T ),
            median3=median(parent_nta_perc_total_time_3, na.rm = T ),
            min3=min(parent_nta_perc_total_time_3, na.rm = T ),
            max3=max(parent_nta_perc_total_time_3, na.rm = T )  ) %>%
  pivot_longer(cols = -group,  
                names_to = c('.value', 'time'), 
                    names_pattern = '(.*)(\\d+)')

write.csv(desctable,"desctable.csv", row.names = TRUE)

```




```{r, echo=FALSE}
desctable %>%
           kbl() %>%
           kable_styling()
```



```{r, echo=FALSE}

# Histograms
# plot
p <- data %>%
  dplyr::select(id,group, parent_nta_perc_total_time_1, parent_nta_perc_total_time_2, parent_nta_perc_total_time_3) %>%
  pivot_longer( cols = starts_with("parent_nta_perc_total_time_"), names_prefix = "time", 
                names_to="t", values_to="outcome") %>%
  mutate(time = ifelse(t == "parent_nta_perc_total_time_1",  1,
                ifelse(t == "parent_nta_perc_total_time_2", 2, 3))) 
  

  
hist<-ggplot(p, aes(x=outcome) )+
  geom_histogram(alpha=0.6, binwidth = 5) +
  xlab("CBCL Externalising") +
  ylab("Frequency") +
  facet_wrap(group~time)

group.labs <- c("PCIT-T", "COS-P", "Waitlist")
names(group.labs) <- c("1", "2", "3")

# New facet label names for group variable
time.labs <- c("Pre", "Post","FollowUp")
names(time.labs) <- c("1", "2","3")
# Create the plot
hist + theme_minimal() +
  facet_grid(
  group ~ time, 
  labeller = labeller(group = group.labs, time = time.labs) )

```


#### Model & Results

A generalized linear mixed model (GLMM) with random effect for individual and fixed effects of group, time and the interaction was run on the number of LP responses in the total number of verbalizations. A binomial distribution was specifed with a logit link. Estimated marginal means were obtained, and contrasts specified to test Hypothesis 1 and Hypothesis 2. Assumptions on the residuals were examined using simulated residuals (see below).


```{r, echo=FALSE}
#  analysis

dat <- data %>%
  dplyr::select(id,group, 
                parent_total_verbalisations_time_1, 
                parent_total_verbalisations_time_2, 
                parent_total_verbalisations_time_3) %>%
  pivot_longer( cols = starts_with("parent_total_verbalisations_time_"), 
                names_prefix = "time", 
                names_to="t", values_to="verbal") %>%
  mutate(time = ifelse(t == "parent_total_verbalisations_time_1",  1,
                ifelse(t == "parent_total_verbalisations_time_2", 2, 3))) 
dat1 <- data %>%
  dplyr::select(id,group, 
                parent_total_nta_time_1,
                parent_total_nta_time_2,
                parent_total_nta_time_3) %>%
  pivot_longer( cols = starts_with("parent_total_nta_time_"), 
                names_prefix = "time", 
                names_to="t", values_to="nta") %>%
  mutate(time = ifelse(t == "parent_total_nta_time_1",  1,
                ifelse(t == "parent_total_nta_time_2", 2, 3))) 


p <- left_join(x=dat,y=dat1, by=c("id","time","group")) 

p$time<- as.factor(p$time)
p$group<- as.factor(p$group)


result0<-glmer(cbind(nta, verbal-nta) ~ (1 | id), data=p, family=binomial)
result1<-glmer(cbind(nta, verbal-nta) ~ group + (1 | id), data=p, family=binomial)
result2<-glmer(cbind(nta, verbal-nta) ~ group + time + (1 | id), data=p, family=binomial)
result3<-glmer(cbind(nta, verbal-nta) ~ group + time +group*time + (1 | id), data=p, family=binomial)

```


This is the ANOVA table for the model. p-values are likelihood ratio tests for adding the new term to the model just before. A p-value <0.05 indicates that adding that term results in better model fit.

```{r, echo=FALSE}
anovatab<-anova(result0,result1,result2,result3)

anovatab <-anovatab %>%
           mutate(Effect = ifelse(npar == 2, "Intercept",
                           ifelse(npar == 4, "Group",
                           ifelse(npar == 6, "Time",
                           ifelse(npar == 9, "Group*Time",""))))) %>%
           relocate(Effect) 
anovatab %>%
           kbl() %>%
           kable_styling()
```



```{r}
summary(result3)
```



#### Estimated Means & Change

Estimated means for each Treatment and Time are presented. In addition, within-group
change between time points is presented with 95% CIs.


```{r, warning=FALSE}
# type="response" means that the emms are transformed and presented on the orginal scale
# This website has a nice demonstration of emmeans with GLMM https://cran.r-project.org/web/packages/emmeans/vignettes/transformations.html

result.change  <- emmeans(result3, "time", by = "group", type = "response")
# Differences give ODDS RATIOS
result.change.d <-pairs(result.change)
summary(result.change)
confint(result.change.d)


```



```{r, echo=FALSE}
result.emm <- emmeans(result3, ~group * time , type = "response")
result.means <- as.data.frame(result.emm)
result.means <-  
  result.means %>%
  mutate(time1= ifelse(group==1 & time==1,  .9,   ## jitter on x axis
                ifelse(group==1 & time==2, 1.9,       
                ifelse(group==1 & time==3, 2.9,       
                ifelse(group==2 & time==1, 1.0,
                ifelse(group==2 & time==2, 2.0,       
                ifelse(group==2 & time==3, 3.0,
                ifelse(group==3 & time==1, 1.1,
                ifelse(group==3 & time==2, 2.1,       
                3.1)))))))))


lp1 <- ggplot(data=result.means, aes(x=time1, y=prob, group=group, shape=group, colour=group)) + geom_line(aes(linetype=group)) + geom_point(size=2) +
    geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL, color=group)) +
  #must define these the same so that only 1 legend is shown
  scale_colour_discrete(name  ="Group",
                            breaks=c(1,2,3),
                            labels=c("PCIT-T", "CoS-P","Waitlist")) +
  scale_shape_discrete(name  ="Group",
                            breaks=c(1,2,3),
                            labels=c("PCIT-T", "CoS-P","Waitlist")) +   
  scale_linetype_manual(name="Group",
                      values =c("solid","dashed","dotted"),
                      breaks=c(1,2,3),
                      labels=c("PCIT-T", "CoS-P","Waitlist")) +
scale_x_continuous("Measurement",breaks = c(1, 2, 3)) +
  scale_y_continuous("Estimated Means (95% CI)") +
theme_classic() 

lp1
  
```


####  Hypothesis 1

We hypothesize that from baseline (‘T1’) to post-treatment/waitlist (‘T2’) there will be a gradient effect, whereby parent-child dyads who receive PCIT-T
will show the greatest gains and parent-child dyads who
receive no treatment will show the least gains. For the outcome, then, the 
null hypothesis is:

$H_0:  \text {mean change}(\text{PCIT-T}) = \text {mean change}(\text{CoS-P}) = \text {mean change}(\text{Waitlist})$
$H_A: \text {mean change}(\text{PCIT-T}) > \text {mean change}(\text{CoS-P}) > \text {mean change}(\text{Waitlist})$


This is a joint hypothesis test.

First, confirming that the orderig of the means is as hypothesized:

```{r, echo=FALSE}
confint(result.change.d)
# df= Inf is ok
```

so, change from t1 to t2 is represented as an OR.  They way these results are presnted, a value >1 means that there is an *decrease* from T1 to T2.  So, for PCIT-T, the OR=3.953, indicating that the odds of a NTA action in the total verbalisations is much larger at Time 1 by a factor of 3.953.  Alternatively, we can take the reciprocal

$\frac{1}{3.953} = 0.25297$

and the CI limits:

$\frac{1}{5.18} = 0.1930502$

$\frac{1}{3.016} = 0.331565$

Here, the odds of an NTA verbalisation are .25 times the odds at time 2 than time 1 for PCIT-T.  Or, there is a 75% decrease in the odds of a NTA verbalisation at Time 2.

Now, to get the p-value associated with the joint hypothesis.

```{r}
## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp1   <- matrix(c(0, 0,0, 0,0, -1,0,0 ),1)
# Hyp2 is 1 df test, 
hyp2   <- matrix(c(0, 0,0, 0,1, 0,1,0 ),1)


# hypF is combined 2df test. Yes!
h<-rbind(hyp1,hyp2)
hypF <- as.matrix(h)
GlobalF <- glht(result3, linfct = hypF, test = Ftest())
GlobalF$linfct
summary(GlobalF, test = Ftest())

#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html

```

p<0.05 so this hypothesis is supported. We have evidence to reject the null in favour of the alternative.


#### Hypothesis 2

There will be a gradient of size of change from T2 to T3 such that the change in the PCIT-T group > the change in the CoS-P group.  This isn't really a gradient; it's just 1 hypothesis.

First, the change in PCITT and COSP from time 2 to time 3:

```{r, echo=FALSE}
result.change
confint(result.change.d)
```




```{r}
## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, -1,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2)
summary(hyp2F)
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html

```

p<0.05,so we rejct the null and have evidence to suggest that the proportions of NTA at time 3 is different for PCITT and COSP.

#### Residuals


Before interpretation, we should examine model assumptions.

For this model (A generalized linear mixed model with binomial distribution and logit link, with random effect for individual), we have to use a different residual than for a linear mixed model. These are simulated residuals, using the DHARMa package in R (citation below).

Under H0 (perfect model, no violations of assumptions), we would expect a straight line in the QQplot. Also, the boxes in the boxplot should range homogenously from 0.25-0.75. To see whether there are deviations from this expectation, the plot calculates a test for uniformity per box, and a test for homgeneity of variances between boxes. 

There are no violations of assumptions.

```{r}

library(DHARMa)

simulationOutput <- simulateResiduals(fittedModel = result3, plot = F)

plot(simulationOutput)


citation("DHARMa")
```



```{r}
rm(list = ls())
```