---
title: "PCIT-CoS-Waitlist RCT Paper 1: BITSEA Competence"
author: "Nancy Briggs"
date: "06 June 2022"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---


### BITSEA Competence {.tabset .tabset-fade .tabset-pills}

Output run on `r Sys.Date()`

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=1)
library(lme4)

library(tidyverse)
library(tidyselect)
library(SASmarkdown)
library(nlme)
library(emmeans)
library(multcomp)
library(kableExtra)
library(ggplot2)
library(lmerTest)
library(effectsize)
library(boot)

options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)

```


```{r, echo=FALSE}
# Data 
# Import data
# recoding for CBCL and totals done in SPSS before bringing data in to R

data <- read.csv("../data_prep.csv", fileEncoding="UTF-8-BOM")
names(data) <- tolower(names(data))

# keep relevant demographic and outcome variables

data <- data %>%
  dplyr::select(id:treatment_time_mins, dobchild, sexchild, numbhousehold, incfamily,
                #DPICS Labelled Praise
                parent_lp_perc_total_time_1, parent_lp_perc_total_time_2, parent_lp_perc_total_time_3,
                #DPICS Negative Talk
                parent_nta_perc_total_time_1, parent_nta_perc_total_time_2, parent_nta_perc_total_time_3,
                #BCAP
                prebcap_risk, post_bcap_risk, fu_bcap_risk,
                #BCAP Lie scale
                prebcap_lie, post_bcap_lie, fu_bcap_lie,
                
                # Child Ext Behav
                cbclext_pre_total,cbclext_post_total,cbclext_fu_total,

                # CBCL Internalising Behav
                cbclint_pre_total,cbclint_post_total,cbclint_fu_total,
                
                #Child Compliance
                total_co_perc_time_1,total_co_perc_time_2,total_co_perc_time_3,
                
                # others, but not essential for paper
                #3. Child Temperament (ECBQ)
                #4. Child social-emotional functioning - 
                #5. Child social-emotional functioning - #2: Child Devereux Early Childhood Assessment (DECA)
                
                #CQQ self Effic.
                pre_ccq_a_topse_mean, post_ccq_a_topse_mean ,fu_ccq_a_topse_mean, 

                # CCQ Caregiving Helplessness
                pre_chqd_sum, post_chqd_sum, fu_chqd_sum,
                
                # CCQ Hostil Parenting
                pre_ccqc_sum, post_ccqc_sum, fu_ccqc_sum,
                
                #1: Brief Infant Toddler Social Emotional Assessment (BITSEA)
                # Problem
                pre_bitseaprob_sum, post_bitseaprob_sum, fu_bitseaprob_sum,
                # Competence
                pre_bitseacomp_sum, post_bitseacomp_sum,fu_bitseacomp_sum

                )


```




#### Descriptives

```{r, echo=FALSE}
# pre_bitseacomp_sum, post_bitseacomp_sum,fu_bitseacomp_sum

desctable <- data %>%
  group_by(group) %>%
  summarise(n1 = sum(!is.na(pre_bitseacomp_sum)),
            mean1=mean(pre_bitseacomp_sum, na.rm = T ),
            sd1=sd(pre_bitseacomp_sum, na.rm = T ),
            median1=median(pre_bitseacomp_sum, na.rm = T ),
            min1=min(pre_bitseacomp_sum, na.rm = T ),
            max1=max(pre_bitseacomp_sum, na.rm = T ),
            
            n2 = sum(!is.na(post_bitseacomp_sum)), 
            mean2=mean(post_bitseacomp_sum, na.rm = T ),
            sd2=sd(post_bitseacomp_sum, na.rm = T ),
            median2=median(post_bitseacomp_sum, na.rm = T ),
            min2=min(post_bitseacomp_sum, na.rm = T ),
            max2=max(post_bitseacomp_sum, na.rm = T ),
            
            n3 = sum(!is.na(fu_bitseacomp_sum)), 
            mean3=mean(fu_bitseacomp_sum, na.rm = T ),
            sd3=sd(fu_bitseacomp_sum, na.rm = T ),
            median3=median(fu_bitseacomp_sum, na.rm = T ),
            min3=min(fu_bitseacomp_sum, na.rm = T ),
            max3=max(fu_bitseacomp_sum, na.rm = T )  ) %>%
  pivot_longer(cols = -group,  
                names_to = c('.value', 'time'), 
                    names_pattern = '(.*)(\\d+)')

write.csv(desctable,"desctable.csv", row.names = TRUE)

```

```{r, echo=FALSE}
desctable %>%
           kbl(digits=3) %>%
           kable_minimal()
```



```{r, echo=FALSE}
##   pre_bitseacomp_mean, post_bitseacomp_mean, fu_bitseacomp_mean,
# Histograms
# plot
p <- data %>%
  dplyr::select(id,group, pre_bitseacomp_sum, post_bitseacomp_sum, fu_bitseacomp_sum) %>%
  pivot_longer( cols = c("pre_bitseacomp_sum", "post_bitseacomp_sum", "fu_bitseacomp_sum"),  names_prefix = "bitseacomp_", 
                names_to="t", values_to="outcome") %>%
  mutate(time = ifelse(t == "pre_bitseacomp_sum",  1,
                ifelse(t == "post_bitseacomp_sum", 2, 3))) 
  

  
hist<-ggplot(p, aes(x=as.numeric(outcome)) )+
  geom_histogram(alpha=0.6, binwidth = 2) +
  xlab("BITSEA Problem") +
  ylab("Frequency") +
  facet_wrap(group~time)

group.labs <- c("PCIT-T", "COS-P", "Waitlist")
names(group.labs) <- c("1", "2", "3")

# New facet label names for group variable
time.labs <- c("Pre", "Post","FollowUp")
names(time.labs) <- c("1", "2","3")
# Create the plot
hist + theme_minimal() +
  facet_grid(
  group ~ time, 
  labeller = labeller(group = group.labs, time = time.labs) )

```

#### Model & Results

A linear mixed model with random effect for individual and fixed effects of group, time and the interaction was run. Estimated marginal means were obtained, and contrasts specified to test Hypothesis 1 and Hypothesis 2.

```{r, echo=FALSE, warning=FALSE}
#  analysis
dat <- data %>%
  dplyr::select(id,group, pre_bitseacomp_sum, post_bitseacomp_sum, fu_bitseacomp_sum) %>%
  pivot_longer( cols = c("pre_bitseacomp_sum", "post_bitseacomp_sum", "fu_bitseacomp_sum"),  names_prefix = "bitseacomp_", 
                names_to="t", values_to="outcome") %>%
  mutate(time = ifelse(t == "pre_bitseacomp_sum",  1,
                ifelse(t == "post_bitseacomp_sum", 2, 3))) 
  

dat$time<- as.factor(dat$time)
dat$group<- as.factor(dat$group)

result0<-lmer(outcome ~ (1 | id), dat)
result1<-lmer(outcome ~ group + (1 | id), dat)
result2<-lmer(outcome ~ group + time + (1 | id), dat)
result3<-lmer(outcome ~ group*time + (1 | id), dat)

```



```{r, echo=FALSE}
anovatab<-anova(result0,result1,result2,result3)

anovatab <-anovatab %>%
           mutate(Effect = ifelse(npar == 3, "Intercept",
                           ifelse(npar == 5, "Group",
                           ifelse(npar == 7, "Time",
                           ifelse(npar == 10, "Group*Time",""))))) %>%
           relocate(Effect) 
anovatab %>%
           kbl(digits=3) %>%
           kable_styling()


## Extract pvaue for evidence / no evidence
evidence = ifelse(anovatab[4,9]<0.05, "evidence", "no evidence")
```



This is the ANOVA table for the model. p-values are likelihood ratio tests for adding the new term to the model just before. A p-value <0.05 indicates that adding that term results in better model fit. The most important p-value is the interaction p-value. Here, there is `r evidence` that the pattern of change over time is different among the groups (p=`r anovatab[4,9]`).



```{r}
summary(result3)
```



#### Estimated Means & Change

Estimated means for each Treatment and Time are presented. In addition, within-group
change between time points is presented with 95% CIs.


```{r child='bootstrapped_es_CI.Rmd'}
```


#### Hypothesis 1

We hypothesize that from baseline (‘T1’) to post-treatment/waitlist (‘T2’) there will be a gradient effect, whereby parent-child dyads who receive PCIT-T
will show the greatest gains and parent-child dyads who
receive no treatment will show the least gains. For the outcome, then, the 
null hypothesis is:

$H_0:  \text {mean change}(\text{PCIT-T}) = \text {mean change}(\text{CoS-P}) = \text {mean change}(\text{Waitlist})$
$H_A: \text {mean change}(\text{PCIT-T}) > \text {mean change}(\text{CoS-P}) > \text {mean change}(\text{Waitlist})$

This is a joint hypothesis test.

First, confirming that the ordering of the means is as hypothesized:

```{r, echo=FALSE}
result.change  <- emmeans(result3, "time", by = "group")
result.change.d <-pairs(result.change)
result.change.d
result.change.d1<-as.data.frame(result.change.d)
result.change.d1<- result.change.d1 %>%
  mutate(t1t2change=-1*estimate) %>%
  mutate(quality=ifelse(t1t2change<0, "decreased t1 to t2", 
                 ifelse(t1t2change>0, "increase t1 to t2", "did not change" )))

result.change.d1 %>%
  filter(contrast=="1 - 2") %>%
  dplyr::select(contrast, group, t1t2change, quality, p.value ) %>%
  kbl(digits=3) %>%
  kable_styling()


# ordering change.
## Extract pvalue for evidence / no evidence
orderincrease <- ifelse( result.change.d1$increase[1]>result.change.d1$increase[4],
         ifelse( result.change.d1$increase[4]>result.change.d1$increase[7] , 
                 "PCITT > COS-P > Waitlist", "Order not satisfied" ))
# orderdecrease <- ifelse( result.change.d1$estimate[1]<result.change.d1$estimate[4],
#          ifelse( result.change.d1$estimate[4]<result.change.d1$estimate[7] , 
#                  "PCITT > COS-P > Waitlist", "Order not satisfied" ))



```

Is the ordering of change satisfied? 

The magnitude of change is 

Greatest to least change from t1 to t2:  `r orderincrease`

Note that I interpreted an increase in scores as being "better".  If a decrease in scores is "better" then these change values should be interpreted in that light.



```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp1   <- matrix(c(0, 0,0, 0,0, -1,0,0 ),1)
# Hyp2 is 1 df test, 
hyp2   <- matrix(c(0, 0,0, 0,1, 0,1,0 ),1)



# hypF is combined 2df test. Yes!
h<-rbind(hyp1,hyp2)
hypF <- as.matrix(h)
GlobalF <- glht(result3, linfct = hypF, test = Ftest())
GlobalF$linfct
summary(GlobalF, test = Ftest())
F<-summary(GlobalF, test = Ftest())

#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html


## Extract pvalue for evidence / no evidence
evidence = ifelse(F$test$pvalue<0.05, "evidence", "no evidence")

```

p= `r round(F$test$pvalue,3)`. We have `r evidence` that there is a difference in the change over from Baseline to Post amongst the groups.




#### Hypothesis 2

There will be a gradient of size of change from T2 to T3 such that the change in the PCIT-T group > the change in the CoS-P group.  This isn't really a gradient; it's just 1 hypothesis.

First, the change in PCITT and COSP from time 2 to time 3:

```{r, echo=FALSE}
result.change.d

# ordering change.
## Extract pvalue for evidence / no evidence
orderincrease <- ifelse( result.change.d1$increase[1]>result.change.d1$increase[4],
                 "PCITT > COS-P, increase was greater in PCITT", 
                 "PCITT < COS-P, increase was not greater in PCITT" )
orderdecrease <- ifelse( result.change.d1$estimate[1]<result.change.d1$estimate[4],
                 "PCITT > COS-P, decrease was greater in PCITT", 
                 "PCITT < COS-P, decrease was not greater in PCITT" )

```

Assuming an increase in scores is better, `r orderincrease`

Assuming a decrease in scores is better,`r orderdecrease`


```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, -1,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2)
summary(hyp2F)
F<-summary(hyp2F)
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html

## Extract pvaue for evidence / no evidence
evidence = ifelse(F$test$pvalue<0.05, "evidence", "no evidence")
```

p>`r round(F$test$pvalue,3)`, so we have `r evidence` to suggest that the change from t2 to t3 in PCITT is greater than the change in COSP.




#### Residuals

Assumptions of the model are met.

```{r, echo=FALSE, warning=FALSE}
plot(result3,form = resid(., type = "pearson") ~ predict(., re.form = NA) )


qqnorm(residuals(result3))
qqline(residuals(result3), col="red")

# plot residuals against predictors
dat <- na.omit(dat)
ggplot(data.frame(x1=dat$group,pearson=residuals(result3,type="pearson")),
       aes(x=x1,y=pearson)) +
  geom_point() +
  theme_bw()

ggplot(data.frame(x2=dat$time,pearson=residuals(result3,type="pearson")),
       aes(x=x2,y=pearson)) +
  geom_point() +
  theme_bw()


```

```{r}
rm(list = ls())
```
