---
title: "PCIT-CoS-Waitlist RCT Paper 1:  Bootstrapping ES CIs"
author: "Nancy Briggs"
date: "05 JAN 2021"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---



##### Group ORs at each timepoint

```{r, echo=FALSE}
result.mean  <- emmeans(result3, "group", by = "time", type="response")
result.mean.d <-pairs(result.mean)
result.mean.d.adj <-pairs(result.mean, adjust="bonferroni")

# save summary of differences as dataset
s_groupdiffs<-summary(result.mean.d)
ci<-confint(result.mean.d)
s_groupdiffsadj<-summary(result.mean.d.adj)
s_groupdiffsadj <- s_groupdiffsadj %>%
        rename(p.value.adj=p.value)

s_groupdiffs<- left_join(s_groupdiffs,s_groupdiffsadj)
s_groupdiffs<- left_join(s_groupdiffs,ci)


```


```{r, echo=FALSE, message=FALSE, include=FALSE, eval=FALSE}
# do not need to bootstrap be OR is the effect size

# bootstrapped percentile CIs for Effect sizes.
fit_mixed <- glmer(outcome ~ group*time + (1 | id), family=binomial, dat)

boot_contrasts <- function(mod){
 # code for ucntion to get group differences at each time point
rg <- ref_grid(mod, mult.levs = rm_levels)

# get the expected means:
em_ <- emmeans(rg, ~ group*time)

# run pairwise tests between the treatment groups within each phase
c_ <- contrast(em_, "pairwise", by = 'time', response)
es_names <- c("t1_1v2", "t1_1v3", "t1_2v3",
              "t2_1v2", "t2_1v3", "t2_2v3")
diffs <- summary(c_)$odds.ratio
names(diffs) <- es_names

# run pairwise tests between the treatment groups within each phase
change_ <- contrast(em_, "pairwise", by = 'group', type="response")
ch_names <- c("g1_1v2",
              "g2_1v2")
change <- summary(change_)$odds.ratio
names(change) <- ch_names  

return(diffs)

}
boot_contrasts_ch <- function(mod){
  rg <- ref_grid(mod, mult.levs = rm_levels)

# get the expected means:
em_ <- emmeans(rg, ~ group*time)

# run pairwise tests between the treatment groups within each phase
c_ <- contrast(em_, "pairwise", by = 'time', type="response")
es_names <- c("t1_1v2", "t1_1v3", "t1_2v3",
              "t2_1v2", "t2_1v3", "t2_2v3")
diffs <- summary(c_)$odds.ratio
names(diffs) <- es_names

# run pairwise tests between the treatment groups within each phase
change_ <- contrast(em_, "pairwise", by = 'group', type="response")
ch_names <- c("g1_1v2"
              "g2_1v2"
              "g3_1v2")
change <- summary(change_)$odds.ratio
names(change) <- ch_names  

return(change)

}
boot_es_group <- function(mod){
  # get baseline pooled SD
  d <- data.frame(outcome= mod@resp$y , mod@pp$X)
  t1_1 <- d %>%
    filter(group2==0 & group3==0 & time2==0 & time3==0 & group2.time2==0  
           & group3.time2==0  & group2.time3==0 ) %>%
    dplyr::select(outcome) %>%
    pull(outcome)
  t1_2 <- d %>%
    filter(group2==1 & group3==0 & time2==0 & time3==0 & group2.time2==0  
           & group3.time2==0  & group2.time3==0 ) %>%
    dplyr::select(outcome) %>%
    pull(outcome)
  t1_3 <- d %>%
    filter(group2==0 & group3==1 & time2==0 & time3==0 & group2.time2==0  
           & group3.time2==0  & group2.time3==0 ) %>%
    dplyr::select(outcome) %>%
    pull(outcome)
  pooledsd<- sd_pooled(t1_1, t1_2, t1_3)
}

numsim <-1
seed=100
groupd <-
  bootMer(fit_mixed, boot_contrasts, nsim = numsim, verbose=FALSE, seed=seed) # R = 599 at least
changed <-
  bootMer(fit_mixed, boot_contrasts_ch, nsim = numsim, verbose=FALSE, seed=seed) # R = 599 at least
bootSD_results <- bootMer(fit_mixed, boot_es_group, nsim = numsim, verbose=FALSE, seed=seed) # R = 599 at least

## extract the bootstrapped values as a data frame ... group differences
bootdiffs<- as.data.frame(groupd)
## extract the bootstrapped values as a data frame ... change differences
bootchanges<- as.data.frame(changed)
## extract the bootstrapped values as a data frame ... pooledSD
bootSDs<- as.data.frame(bootSD_results)
bootdiffs <- cbind(bootdiffs,bootchanges,bootSDs)
#reshape long;
es <- bootdiffs %>%
  pivot_longer( cols = starts_with("t"),  
                names_to="comparison", values_to="diff") %>%
    pivot_longer( cols = starts_with("g"),  
                names_to="comparison2", values_to="change") %>%
  mutate(es=diff/V1) %>%
  mutate(esch=change/V1) %>%
  group_by(comparison)   %>%
  summarise(es2_5  = ( quantile(es, c(0.025), na.rm=TRUE)),
            es97_5 = quantile(es, c(0.975), na.rm=TRUE),
            esch2_5  = ( quantile(esch, c(0.025), na.rm=TRUE)),
            esch97_5 = quantile(esch, c(0.975), na.rm=TRUE))


es_gpdiffs <- bootdiffs %>%
  pivot_longer( cols = starts_with("t"),  
                names_to="comparison", values_to="diff") %>%
  mutate(es=diff/V1) %>%
  group_by(comparison)   %>%
  summarise(es2_5  = ( quantile(es, c(0.025), na.rm=TRUE)) ,
            es97_5 = quantile(es, c(0.975), na.rm=TRUE) )

es_change <- bootdiffs %>%
    pivot_longer( cols = starts_with("g"),  
                names_to="comparison", values_to="change") %>%
  mutate(esch=change/V1) %>%
  group_by(comparison)   %>%
  summarise(esch2_5  = ( quantile(esch, c(0.025), na.rm=TRUE)),
            esch97_5 = quantile(esch, c(0.975), na.rm=TRUE))


s_groupdiffs<- cbind(s_groupdiffs,es_gpdiffs)
```


```{r, echo=FALSE}
s_groupdiffs %>%
  dplyr::select(!c(df, null)) %>%
  kbl(digits=3,caption = "Odds Ratios at Each Time Point", 
  col.names = c("Comparison", "Time","OR", "SE", "z","p","adj p",
                  "Lower CL","Upper CL" ))  %>%
  kable_styling()
```

In this table the OR is the effect size.  No bootstrapping is needed.



At t1, the OR for G1 vs G2 is `r round(s_groupdiffs$odds.ratio[1],3)`.

At t1, the OR for G1 vs G3 is `r round(s_groupdiffs$odds.ratio[2],3)`.

At t1, the OR for G2 vs G3 is `r round(s_groupdiffs$odds.ratio[3],3)`.

None of the pvalues < 0.05.

##### Change within each group

```{r, echo=FALSE, warning=FALSE}

result.change  <- emmeans(result3, "time", by = "group",
                          type="response", reverse=TRUE)
result.change.d <-pairs(result.change, reverse=TRUE)
result.change.d.adj <-pairs(result.change,reverse=TRUE, adjust="bonferroni")

# save summary of differences as dataset
s<-summary(result.change.d)
ci<-confint(result.change.d)
sadj<-summary(result.change.d.adj)
sadj <- sadj %>%
        rename(p.value.adj=p.value)

s<- left_join(s,sadj)
s<- left_join(s,ci)

s %>%
  dplyr::select(c(-null,-df)) %>%
  kbl(digits=3,caption = "Within-Group Change", 
  col.names = c("Comparison", "Group","OR", "SE", "z","p","adj p",
                  "Lower CL","Upper CL") )  %>%
  kable_styling() 
```

These ORs now represent the odds of a 1 at time 2 relative to time 1 within group. Because there is only 1 comparison within group, the adjusted pvalue is the same.


Now to test whether the 2 above ORs are different:

```{r, echo=F}



```


```{r, echo=FALSE, warning=FALSE}

result.emm <- emmeans(result3, ~group * time, type="response")
result.means <- as.data.frame(result.emm)
result.means <-  
  result.means %>%
  mutate(time1= ifelse(group==1 & time==1,  .9,   ## jitter on x axis
                ifelse(group==1 & time==2, 1.9,       
                ifelse(group==1 & time==3, 2.9,       
                ifelse(group==2 & time==1, 1.0,
                ifelse(group==2 & time==2, 2.0,       
                ifelse(group==2 & time==3, 3.0,
                ifelse(group==3 & time==1, 1.1,
                ifelse(group==3 & time==2, 2.1,       
                3.1)))))))))


lp1 <- ggplot(data=result.means, aes(x=time1, y=prob, group=group, shape=group, colour=group)) + geom_line(aes(linetype=group)) + geom_point(size=2) +
    geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL, color=group)) +
  #must define these the same so that only 1 legend is shown
  scale_colour_discrete(name  ="Group",
                            breaks=c(1,2,3),
                            labels=c("PCIT-T", "CoS-P","Waitlist")) +
  scale_shape_discrete(name  ="Group",
                            breaks=c(1,2,3),
                            labels=c("PCIT-T", "CoS-P","Waitlist")) +   
  scale_linetype_manual(name="Group",
                      values =c("solid","dashed","dotted"),
                      breaks=c(1,2,3),
                      labels=c("PCIT-T", "CoS-P","Waitlist")) +
scale_x_continuous("Measurement",breaks = c(1, 2, 3)) +
  scale_y_continuous("Estimated Probability (95% CI)") +
theme_classic() 

lp1
  
```
