---
title: "PCIT-CoS-Waitlist RCT Paper 1: BCAP, Lie Score (Time Varying)"
author: "Nancy Briggs"
date: "05 JAN 2021"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---


### BCAP {.tabset .tabset-fade .tabset-pills}

This is an analysis of BCAP, with the time-varying covariate of lie included in the model. I did not include tests of Hypothesis 1 & 2, as I am not sure how to do that in the presence of a lie*group *time interaction.

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)

library(tidyverse)
library(tidyselect)
library(SASmarkdown)
library(nlme)
library(emmeans)
library(multcomp)
library(kableExtra)
library(ggplot2)
library(lmerTest)
library(effectsize)
library(boot)

options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)

```

```{r, echo=FALSE}
# Data 
# Import data
# recoding for CBCL and totals done in SPSS before bringing data in to R

data <- read.csv("../data_prep.csv", fileEncoding="UTF-8-BOM")
names(data) <- tolower(names(data))

# keep relevant demographic and outcome variables

data <- data %>%
  dplyr::select(id:treatment_time_mins, dobchild, sexchild, numbhousehold, incfamily,
                #DPICS Labelled Praise
                parent_lp_perc_total_time_1, parent_lp_perc_total_time_2, parent_lp_perc_total_time_3,
                #DPICS Negative Talk
                parent_nta_perc_total_time_1, parent_nta_perc_total_time_2, parent_nta_perc_total_time_3,
                #BCAP
                prebcap_risk, post_bcap_risk, fu_bcap_risk,
                #BCAP Lie scale
                prebcap_lie, post_bcap_lie, fu_bcap_lie,
                
                # Child Ext Behav
                cbclext_pre_total,cbclext_post_total,cbclext_fu_total,

                # CBCL Internalising Behav
                cbclint_pre_total,cbclint_post_total,cbclint_fu_total,
                
                #Child Compliance
                total_co_perc_time_1,total_co_perc_time_2,total_co_perc_time_3
                
                # others, but not essential for paper
                #3. Child Temperament (ECBQ)
                #4. Child social-emotional functioning - #1: Brief Infant Toddler Social Emotional Assessment (BITSEA)
                #5. Child social-emotional functioning - #2: Child Devereux Early Childhood Assessment (DECA)
                
                )


```




#### Descriptives

```{r, echo=FALSE}
 #BCAP  prebcap_risk, post_bcap_risk, fu_bcap_risk,
#BCAP Lie scale prebcap_lie, post_bcap_lie, fu_bcap_lie,

data$prebcap_risk<-as.double(data$prebcap_risk)
data$post_bcap_risk<-as.double(data$post_bcap_risk)
data$fu_bcap_risk<-as.double(data$fu_bcap_risk)
data$prebcap_lie<-as.double(data$prebcap_lie)
data$post_bcap_lie<-as.double(data$post_bcap_lie)
data$fu_bcap_lie<-as.double(data$fu_bcap_lie)

desctable <- data %>%
  group_by(group) %>%
  rename(bcaprisk_pre = prebcap_risk) %>%
  rename(bcaprisk_post = post_bcap_risk) %>%
  rename(bcaprisk_fu = fu_bcap_risk) %>%
  mutate(score1 = case_when( 
    bcaprisk_pre == -9 ~ NA_real_,       #assign missing values to -9
    TRUE ~ bcaprisk_pre)) %>%
  mutate(score2 = case_when( 
    bcaprisk_post == -9 ~ NA_real_,   #assign missing values to -9
    TRUE ~ bcaprisk_post)) %>%
  mutate(score3 = case_when( 
    bcaprisk_fu == -9 ~ NA_real_,     #assign missing values to -9
    TRUE ~ bcaprisk_fu)) %>%
  summarise(n1 = sum(!is.na(score1)) , 
            mean1=mean(cbclint_pre_total, na.rm = T ),
            sd1=sd(cbclint_pre_total, na.rm = T ),
            median1=median(cbclint_pre_total, na.rm = T ),
            min1=min(cbclint_pre_total, na.rm = T ),
            max1=max(cbclint_pre_total, na.rm = T ),
            
            n2 = sum(!is.na(score2)) , 
            mean2=mean(cbclint_post_total, na.rm = T ),
            sd2=sd(cbclint_post_total, na.rm = T ),
            median2=median(cbclint_post_total, na.rm = T ),
            min2=min(cbclint_post_total, na.rm = T ),
            max2=max(cbclint_post_total, na.rm = T ),
            
            n3 = sum(!is.na(score3)) , 
            mean3=mean(cbclint_fu_total, na.rm = T ),
            sd3=sd(cbclint_fu_total, na.rm = T ),
            median3=median(cbclint_fu_total, na.rm = T ),
            min3=min(cbclint_fu_total, na.rm = T ),
            max3=max(cbclint_fu_total, na.rm = T )  ) %>%
  pivot_longer(cols = -group,  
                names_to = c('.value', 'time'), 
                    names_pattern = '(.*)(\\d+)')

write.csv(desctable,"desctable.csv", row.names = TRUE)


desctable_lie <- data %>%
  group_by(group) %>%
  rename(bcaplie_pre = prebcap_lie) %>%
  rename(bcaplie_post = post_bcap_lie) %>%
  rename(bcaplie_fu = fu_bcap_lie) %>%
  mutate(score1 = case_when( 
    bcaplie_pre == -9 ~ NA_real_,     #assign missing values to zeros
    TRUE ~ bcaplie_pre)) %>%
  mutate(score2 = case_when( 
    bcaplie_post == -9 ~ NA_real_,     #assign missing values to zeros
    TRUE ~ bcaplie_post)) %>%
  mutate(score3 = case_when( 
    bcaplie_fu == -9 ~ NA_real_,     #assign missing values to zeros
    TRUE ~ bcaplie_fu)) %>%
  summarise(n1 = sum(!is.na(score1)) , 
            mean1=mean(bcaplie_pre, na.rm = T ),
            sd1=sd(bcaplie_pre, na.rm = T ),
            median1=median(bcaplie_pre, na.rm = T ),
            min1=min(bcaplie_pre, na.rm = T ),
            max1=max(bcaplie_pre, na.rm = T ),
            
            n2 = sum(!is.na(score2)) , 
            mean2=mean(bcaplie_post, na.rm = T ),
            sd2=sd(bcaplie_post, na.rm = T ),
            median2=median(bcaplie_post, na.rm = T ),
            min2=min(bcaplie_post, na.rm = T ),
            max2=max(bcaplie_post, na.rm = T ),
            
            n3 = sum(!is.na(score3)) , 
            mean3=mean(bcaplie_fu, na.rm = T ),
            sd3=sd(bcaplie_fu, na.rm = T ),
            median3=median(bcaplie_fu, na.rm = T ),
            min3=min(bcaplie_fu, na.rm = T ),
            max3=max(bcaplie_fu, na.rm = T )  ) %>%
  pivot_longer(cols = -group,  
                names_to = c('.value', 'time'), 
                    names_pattern = '(.*)(\\d+)')

write.csv(desctable_lie,"desctable_lie.csv", row.names = TRUE)

```
```{r, echo=FALSE}
#BCAP  prebcap_risk, post_bcap_risk, fu_bcap_risk,
#BCAP Lie scale prebcap_lie, post_bcap_lie, fu_bcap_lie,




data$prebcap_risk<-as.double(data$prebcap_risk)
data$post_bcap_risk<-as.double(data$post_bcap_risk)
data$fu_bcap_risk<-as.double(data$fu_bcap_risk)
data$prebcap_lie<-as.double(data$prebcap_lie)
data$post_bcap_lie<-as.double(data$post_bcap_lie)
data$fu_bcap_lie<-as.double(data$fu_bcap_lie)

desctable_nomissing <- data %>%
  group_by(group) %>%
  rename(bcaprisk_pre = prebcap_risk) %>%
  rename(bcaprisk_post = post_bcap_risk) %>%
  rename(bcaprisk_fu = fu_bcap_risk) %>%
  mutate(score1 = case_when( 
    bcaprisk_pre == -9 ~ NA_real_,       #assign missing values to -9
    TRUE ~ bcaprisk_pre)) %>%
  mutate(score2 = case_when( 
    bcaprisk_post == -9 ~ NA_real_,   #assign missing values to -9
    TRUE ~ bcaprisk_post)) %>%
  mutate(score3 = case_when( 
    bcaprisk_fu == -9 ~ NA_real_,     #assign missing values to -9
    TRUE ~ bcaprisk_fu)) 

desctable_nomissing$bcaprisk_pre[is.na(desctable_nomissing$prebcap_lie)]<-NA
desctable_nomissing$bcaprisk_post[is.na(desctable_nomissing$post_bcap_lie)]<-NA
desctable_nomissing$bcaprisk_fu[is.na(desctable_nomissing$fu_bcap_lie)]<-NA

desctable_nomissing$score1[is.na(desctable_nomissing$prebcap_lie)]<-NA
desctable_nomissing$score2[is.na(desctable_nomissing$post_bcap_lie)]<-NA
desctable_nomissing$score3[is.na(desctable_nomissing$fu_bcap_lie)]<-NA


desctable_nomissing <- desctable_nomissing %>%
  group_by(group) %>%
  summarise(n1 = sum(!is.na(score1)) , 
            mean1=mean(cbclint_pre_total, na.rm = T ),
            sd1=sd(cbclint_pre_total, na.rm = T ),
            median1=median(cbclint_pre_total, na.rm = T ),
            min1=min(cbclint_pre_total, na.rm = T ),
            max1=max(cbclint_pre_total, na.rm = T ),
            
            n2 = sum(!is.na(score2)) , 
            mean2=mean(cbclint_post_total, na.rm = T ),
            sd2=sd(cbclint_post_total, na.rm = T ),
            median2=median(cbclint_post_total, na.rm = T ),
            min2=min(cbclint_post_total, na.rm = T ),
            max2=max(cbclint_post_total, na.rm = T ),
            
            n3 = sum(!is.na(score3)) , 
            mean3=mean(cbclint_fu_total, na.rm = T ),
            sd3=sd(cbclint_fu_total, na.rm = T ),
            median3=median(cbclint_fu_total, na.rm = T ),
            min3=min(cbclint_fu_total, na.rm = T ),
            max3=max(cbclint_fu_total, na.rm = T )  ) %>%
  pivot_longer(cols = -group,  
                names_to = c('.value', 'time'), 
                    names_pattern = '(.*)(\\d+)')

write.csv(desctable,"desctable_nomissing.csv", row.names = TRUE)





data$prebcap_risk<-as.double(data$prebcap_risk)
data$post_bcap_risk<-as.double(data$post_bcap_risk)
data$fu_bcap_risk<-as.double(data$fu_bcap_risk)
data$prebcap_lie<-as.double(data$prebcap_lie)
data$post_bcap_lie<-as.double(data$post_bcap_lie)
data$fu_bcap_lie<-as.double(data$fu_bcap_lie)


desctable_nomissing_lie <- data %>%
  group_by(group) %>%
  rename(bcaplie_pre = prebcap_lie) %>%
  rename(bcaplie_post = post_bcap_lie) %>%
  rename(bcaplie_fu = fu_bcap_lie) %>%
  mutate(score1 = case_when( 
    bcaplie_pre == -9 ~ NA_real_,     #assign missing values to zeros
    TRUE ~ bcaplie_pre)) %>%
  mutate(score2 = case_when( 
    bcaplie_post == -9 ~ NA_real_,     #assign missing values to zeros
    TRUE ~ bcaplie_post)) %>%
  mutate(score3 = case_when( 
    bcaplie_fu == -9 ~ NA_real_,     #assign missing values to zeros
    TRUE ~ bcaplie_fu))

desctable_nomissing_lie$bcaplie_pre[is.na(desctable_nomissing_lie$prebcap_risk)]<-NA
desctable_nomissing_lie$bcaplie_post[is.na(desctable_nomissing_lie$post_bcap_risk)]<-NA
desctable_nomissing_lie$bcaplie_fu[is.na(desctable_nomissing_lie$fu_bcap_risk)]<-NA

desctable_nomissing_lie$score1[is.na(desctable_nomissing_lie$prebcap_risk)]<-NA
desctable_nomissing_lie$score2[is.na(desctable_nomissing_lie$post_bcap_risk)]<-NA
desctable_nomissing_lie$score3[is.na(desctable_nomissing_lie$fu_bcap_risk)]<-NA

desctable_nomissing_lie <- desctable_nomissing_lie %>%
  summarise(n1 = sum(!is.na(score1)) , 
            mean1=mean(bcaplie_pre, na.rm = T ),
            sd1=sd(bcaplie_pre, na.rm = T ),
            median1=median(bcaplie_pre, na.rm = T ),
            min1=min(bcaplie_pre, na.rm = T ),
            max1=max(bcaplie_pre, na.rm = T ),
            
            n2 = sum(!is.na(score2)) , 
            mean2=mean(bcaplie_post, na.rm = T ),
            sd2=sd(bcaplie_post, na.rm = T ),
            median2=median(bcaplie_post, na.rm = T ),
            min2=min(bcaplie_post, na.rm = T ),
            max2=max(bcaplie_post, na.rm = T ),
            
            n3 = sum(!is.na(score3)) , 
            mean3=mean(bcaplie_fu, na.rm = T ),
            sd3=sd(bcaplie_fu, na.rm = T ),
            median3=median(bcaplie_fu, na.rm = T ),
            min3=min(bcaplie_fu, na.rm = T ),
            max3=max(bcaplie_fu, na.rm = T )  ) %>%
  pivot_longer(cols = -group,  
                names_to = c('.value', 'time'), 
                    names_pattern = '(.*)(\\d+)')

write.csv(desctable_nomissing_lie,"desctable_nomissing_lie.csv", row.names = TRUE)


```


##### BCAP Risk descriptives
##### All valid BCAP values included
```{r, echo=FALSE}
desctable %>%
           kbl() %>%
           kable_styling()
```



##### BCAP Lie descriptives
```{r, echo=FALSE}
desctable_lie %>%
           kbl() %>%
           kable_styling()
```

##### BCAP values, no missing BACP or Lie values
```{r, echo=FALSE}
desctable_nomissing %>%
           kbl() %>%
           kable_styling()
```

##### BCAP Lie values, no missing BACP or Lie values
```{r, echo=FALSE}
desctable_nomissing_lie %>%
           kbl() %>%
           kable_styling()
```



```{r, echo=FALSE}

# Histograms
# plot
p <- data %>%
    dplyr::select(id,group, 
                  prebcap_risk, post_bcap_risk, fu_bcap_risk, prebcap_lie) %>%
  rename(bcaprisk_pre = prebcap_risk) %>%
  rename(bcaprisk_post = post_bcap_risk) %>%
  rename(bcaprisk_fu = fu_bcap_risk) %>%
  pivot_longer( cols = c(-id,-group, -prebcap_lie), 
                names_prefix = "bcaprisk_", 
                names_to="t", values_to="outcome") %>% 
  mutate(time = ifelse(t == "pre",  1,
                ifelse(t == "post", 2, 3))) 
plie <- data %>%
    dplyr::select(id,group, 
                  prebcap_lie, post_bcap_lie, fu_bcap_lie) %>%
  rename(bcaplie_pre = prebcap_lie) %>%
  rename(bcaplie_post = post_bcap_lie) %>%
  rename(bcaplie_fu = fu_bcap_lie) %>%
  pivot_longer( cols = c(-id,-group), 
                names_prefix = "bcaplie_", 
                names_to="t", values_to="lie") %>% 
  mutate(time = ifelse(t == "pre",  1,
                ifelse(t == "post", 2, 3)))

p <- full_join(x=p,y=plie) 
# need all models fit to the same data  Since some lie values are missing, then need to exclude them.
p$outcome[is.na(p$lie)]<-NA
p<- p %>% drop_na(outcome)


hist0<-ggplot(p, aes(x=outcome) )+
  geom_histogram(alpha=0.6, binwidth = 5) +
  xlab("BCAP Risk") +
  ylab("Frequency") 
hist0 +   theme_minimal()

hist<-ggplot(p, aes(x=outcome) )+
  geom_histogram(alpha=0.6, binwidth = 5) +
  xlab("BCAP Risk") +
  ylab("Frequency") +
  facet_wrap(group~time)

group.labs <- c("PCIT-T", "COS-P", "Waitlist")
names(group.labs) <- c("1", "2", "3")

# New facet label names for group variable
time.labs <- c("Pre", "Post","FollowUp")
names(time.labs) <- c("1", "2","3")
# Create the plot
hist + theme_minimal() +
  facet_grid(
  group ~ time, 
  labeller = labeller(group = group.labs, time = time.labs) )





```

```{r, echo=FALSE, warning=FALSE, message=FALSE}
#Plots to visualise the relationship between TVC (lie) and outcome

p1<- p %>%
  filter(group==1)
# Pre-specify some variables to make the following code more generic
d  <- p1
x  <- 'time'
y1 <- 'outcome'
y2 <- 'lie'

#-----------------------------------------------------------------------------
# Rescale the second y axis by 
#   - subtracting its minimum value (to set it to start at 0)
#   - scaling so that it has the same range as the 'y1' variable
#   - offsettting it by the minimum value of y1
#-----------------------------------------------------------------------------
a            <- range(d[[y1]])
b            <- range(d[[y2]])
scale_factor <- diff(a)/diff(b)
d[[y2]]      <- ((d[[y2]] - b[1]) * scale_factor) + a[1]
#-----------------------------------------------------------------------------
# Need to define the second axis transformation to be the inverse of the data
# transformation to everything cancels out appropriately
#-----------------------------------------------------------------------------
trans <- ~ ((. - a[1]) / scale_factor) + b[1]

#-----------------------------------------------------------------------------
# tell the y axis to set up a scaled secondary axis with the given transform
#-----------------------------------------------------------------------------
ggplot(d) +
  geom_point(aes_string(x, y1)) +
  geom_line(aes_string(x, y1)) + 
  geom_point(aes_string(x, y2), col='red') + 
  geom_line(aes_string(x, y2), col='red') + 
  scale_y_continuous(sec.axis = sec_axis(trans=trans, name=y2)) +
      facet_wrap(~id) + 
  theme(  legend.position="bottom")+
  scale_x_continuous("Measurement",breaks = c(1, 2, 3)) +
   labs(title = "BCAP Risk (black) & Lie (red) Scores",
        subtitle ="PCITT")





p2<- p %>%
  filter(group==2)
# Pre-specify some variables to make the following code more generic
d  <- p2
x  <- 'time'
y1 <- 'outcome'
y2 <- 'lie'
#-----------------------------------------------------------------------------
# Rescale the second y axis by 
#   - subtracting its minimum value (to set it to start at 0)
#   - scaling so that it has the same range as the 'y1' variable
#   - offsettting it by the minimum value of y1
#-----------------------------------------------------------------------------
a            <- range(d[[y1]])
b            <- range(d[[y2]])
scale_factor <- diff(a)/diff(b)
d[[y2]]      <- ((d[[y2]] - b[1]) * scale_factor) + a[1]
#-----------------------------------------------------------------------------
# Need to define the second axis transformation to be the inverse of the data
# transformation to everything cancels out appropriately
#-----------------------------------------------------------------------------
trans <- ~ ((. - a[1]) / scale_factor) + b[1]

#-----------------------------------------------------------------------------
# tell the y axis to set up a scaled secondary axis with the given transform
#-----------------------------------------------------------------------------
ggplot(d) +
  geom_point(aes_string(x, y1)) +
  geom_line(aes_string(x, y1)) + 
  geom_point(aes_string(x, y2), col='red') + 
  geom_line(aes_string(x, y2), col='red') + 
  scale_y_continuous(sec.axis = sec_axis(trans=trans, name=y2)) +
      facet_wrap(~id) + 
  theme(  legend.position="bottom")+
  scale_x_continuous("Measurement",breaks = c(1, 2, 3)) +
   labs(title = "BCAP Risk (black) & Lie (red) Scores",
        subtitle ="COSP")




p3<- p %>%
  filter(group==3)
# Pre-specify some variables to make the following code more generic
d  <- p3
x  <- 'time'
y1 <- 'outcome'
y2 <- 'lie'
#-----------------------------------------------------------------------------
# Rescale the second y axis by 
#   - subtracting its minimum value (to set it to start at 0)
#   - scaling so that it has the same range as the 'y1' variable
#   - offsettting it by the minimum value of y1
#-----------------------------------------------------------------------------
a            <- range(d[[y1]])
b            <- range(d[[y2]])
scale_factor <- diff(a)/diff(b)
d[[y2]]      <- ((d[[y2]] - b[1]) * scale_factor) + a[1]
#-----------------------------------------------------------------------------
# Need to define the second axis transformation to be the inverse of the data
# transformation to everything cancels out appropriately
#-----------------------------------------------------------------------------
trans <- ~ ((. - a[1]) / scale_factor) + b[1]

#-----------------------------------------------------------------------------
# tell the y axis to set up a scaled secondary axis with the given transform
#-----------------------------------------------------------------------------
ggplot(d) +
  geom_point(aes_string(x, y1)) +
  geom_line(aes_string(x, y1)) + 
  geom_point(aes_string(x, y2), col='red') + 
  geom_line(aes_string(x, y2), col='red') + 
  scale_y_continuous(sec.axis = sec_axis(trans=trans, name=y2)) +
      facet_wrap(~id) + 
  theme(  legend.position="bottom")+
  scale_x_continuous("Measurement",breaks = c(1, 2, 3)) +
   labs(title = "BCAP Risk (black) & Lie (red) Scores",
        subtitle ="Waitlist")



```


#### Model & Results

A linear mixed model with random effect for individual and fixed effects of group, time and the interaction was run, and the time-varying covariate of BCAP lie score included as predictor. This results in an analysis sample of N=153. Estimated marginal means were obtained for low, mean and high lie scores, and plots for a person who showed *consistent* low, mean or high lie scores is presented.



```{r, echo=FALSE, warning=FALSE, message=FALSE}
#  analysis
# need all models fit to the same data  Since some lie values are missing, then need to exclude them.

dat <- p 
dat$outcome[is.na(dat$lie)]<-NA
dat<- dat %>% drop_na(outcome)


dat$time<- as.factor(dat$time)
dat$group<- as.factor(dat$group)

result0<-lmer(outcome ~ (1 | id), dat, na.action=na.omit)
result1<-lmer(outcome ~ lie + (1 | id), dat)
result2<-lmer(outcome ~ lie + group + (1 | id), dat)
result3<-lmer(outcome ~ lie + group + time + (1 | id), dat)
result4<-lmer(outcome ~ lie + group*time + lie + (1 | id), dat)
result5<-lmer(outcome ~ lie*group*time + lie + (1 | id), dat)

```




This is the ANOVA table for the model. p-values are likelihood ratio tests for adding the new term to the model just before. A p-value <0.05 indicates that adding that term results in better model fit. As you can see, the model with the lie*group *time interaction showed significant imporement over the model with a min effect of lie and the group*time interaction (p<0.001) .

```{r, echo=FALSE}
anovatab<-anova(result0,result1,result2,result3,result4,result5)

anovatab <-anovatab %>%
           mutate(Effect = ifelse(npar == 3, "Intercept",
                           ifelse(npar == 4, "Lie",
                           ifelse(npar == 6, "Lie + Group",
                           ifelse(npar == 8, "Lie + Group + Time", 
                           ifelse(npar == 11, "Lie + Group* Time", 
                                  "Lie*Group* Time")))))) %>%
           relocate(Effect) 
anovatab %>%
           kbl() %>%
           kable_styling()
```



```{r}
summary(result5)
```



#### Estimated Means & Change

For the three-way interaction, I obtained estimates ofmeans for each group*time combination, at low, mean and high lie scores.  That is, at each time point, I calculated the mean lie score, and a low lie value is 1 sd below the mean for that
timepoint.  Then I plotted the group*time interaction for a person who showed consistent (across all 3 timepoints), low / mean / high values of lie score. Obviously, there is the possibility of people jumping around in their lie score so that, for example, they my show a high value (above the mean at T1), but perhaps show the mean score atT2 and T3.  But I can't make a plot for all possibilities.

Estimated means for each Treatment and Time are presented. In addition, within-group
change between time points is presented with 95% CIs.


```{r, echo=FALSE, warning=FALSE}

# use means of lie scale at each timeppoint to calculate the EMS for risk
lie1 <- mean(dat$lie[dat$time==1])
lie2 <- mean(dat$lie[dat$time==2])
lie3 <- mean(dat$lie[dat$time==3])

result.change  <- emmeans(result5, pairwise ~ time + lie, by = "group", 
                          at = list(lie = c(lie1,lie2,lie3)) )
result.change.d <-pairs(result.change)

# Extract the correctvalues (meaning the EMMS based on the correct lie mean)
means<-transform(result.change, CI.width = upper.CL - lower.CL)
means <- means %>%
         filter(time==1 & lie==lie1 | 
                time==2 & lie==lie2 | 
                time==3 & lie==lie3) %>%
         mutate(lie=as.numeric(lie), 
                group=as.integer(group),
                time=as.numeric(time)) %>%
         dplyr::select(-c("contrast"))  %>%
         mutate(lieTV=1)   ## low lie at each time point

# low lie score
# use means of lie scale at each timepoint to calculate the EMS for risk
lielow1 <- mean(dat$lie[dat$time==1]) - sd(dat$lie[dat$time==1])
lielow2 <- mean(dat$lie[dat$time==2]) - sd(dat$lie[dat$time==2])
lielow3 <- mean(dat$lie[dat$time==3]) - sd(dat$lie[dat$time==3])

result.change.low  <- emmeans(result5, pairwise ~ time + lie, by = "group", 
                          at = list(lie = c(lielow1,lielow2,lielow3)) )
result.change.d.low  <-pairs(result.change.low )

# Extract the correctvalues (meaning the EMMS based on the correct lie mean)
means.low <-transform(result.change.low , CI.width = upper.CL - lower.CL)
means.low  <- means.low  %>%
         filter(time==1 & lie==lielow1 | 
                time==2 & lie==lielow2 | 
                time==3 & lie==lielow3) %>%
         mutate(lie=as.numeric(lie), 
                group=as.integer(group),
                time=as.numeric(time)) %>%
         dplyr::select(-c("contrast")) %>%
         mutate(lieTV=0)   ## mean lie at each time point

# high lie score
# use means of lie scale at each timepoint to calculate the EMS for risk
liehi1 <- mean(dat$lie[dat$time==1]) + sd(dat$lie[dat$time==1])
liehi2 <- mean(dat$lie[dat$time==2]) + sd(dat$lie[dat$time==2])
liehi3 <- mean(dat$lie[dat$time==3]) + sd(dat$lie[dat$time==3])

result.change.hi  <- emmeans(result5, pairwise ~ time + lie, by = "group", 
                          at = list(lie = c(liehi1,liehi2,liehi3)) )
result.change.d.hi  <-pairs(result.change.hi )

# Extract the correctvalues (meaning the EMMS based on the correct lie mean)
means.hi <-transform(result.change.hi , CI.width = upper.CL - lower.CL)
means.hi  <- means.hi  %>%
         filter(time==1 & lie==liehi1 | 
                time==2 & lie==liehi2 | 
                time==3 & lie==liehi3) %>%
         mutate(lie=as.numeric(lie), 
                group=as.integer(group),
                time=as.numeric(time)) %>%
         dplyr::select(-c("contrast")) %>%
         mutate(lieTV=2)   ## high lie at each time point


liemeans <- rbind(means.low, means, means.hi)


```



Means, assuming consistently low, average or high lie scores

```{r, echo=FALSE, warning=FALSE, message=FALSE}
liemeans %>%
           kbl(digits=3) %>%
           kable_styling()


liemeans <-  
  liemeans %>%
  mutate(time1= ifelse(group==1 & time==1,  .9,   ## jitter on x axis
                ifelse(group==1 & time==2, 1.9,       
                ifelse(group==1 & time==3, 2.9,       
                ifelse(group==2 & time==1, 1.0,
                ifelse(group==2 & time==2, 2.0,       
                ifelse(group==2 & time==3, 3.0,
                ifelse(group==3 & time==1, 1.1,
                ifelse(group==3 & time==2, 2.1,       
                3.1)))))))))


liemeans <- liemeans %>%
  dplyr::mutate(group=as.factor(group)) %>%
  dplyr::mutate(lieTV=as.factor(lieTV))

# labels for facet columns (lie TV groups)
liemeans$lieTV <- factor(liemeans$lieTV, levels = c("0", "1", "2"), 
                  labels = c("Low Lie Score", "Mean Lie Score", "High Lie Score"))

Lieplot <- ggplot(data=liemeans, aes(x=time1, y=emmean, group=group, shape=group, colour=group)) + geom_line(aes(linetype=group)) + geom_point(size=2) +
    geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL, color=group)) +
  #must define these the same so that only 1 legend is shown
  scale_colour_discrete(name  ="Group",
                            breaks=c(1,2,3),
                            labels=c("PCIT-T", "CoS-P","Waitlist")) +
  scale_shape_discrete(name  ="Group",
                            breaks=c(1,2,3),
                            labels=c("PCIT-T", "CoS-P","Waitlist")) +   
  scale_linetype_manual(name="Group",
                      values =c("solid","dashed","dotted"),
                      breaks=c(1,2,3),
                      labels=c("PCIT-T", "CoS-P","Waitlist")) +
scale_x_continuous("Measurement",breaks = c(1, 2, 3)) +
  scale_y_continuous("Estimated Means (95% CI)") +
theme_classic() + theme(legend.position = "top") + 
  facet_grid(.~lieTV)

Lieplot


```
It appears that the effect of lie score is especially pronounced at time 1, for the COS-P group. You can see that the estimated mean for the COSP group at time 1 in the high group is very high, and quickly decreases whereas that doesn't happen in the Low or Mean groups, nor is it as pronounced in the other groups.


<!-- #### Hypothesis 1 -->

<!-- We hypothesize that from baseline (‘T1’) to post-treatment/waitlist (‘T2’) there will be a gradient effect, whereby parent-child dyads who receive PCIT-T -->
<!-- will show the greatest gains and parent-child dyads who -->
<!-- receive no treatment will show the least gains. For the outcome, then, the  -->
<!-- null hypothesis is: -->

<!-- $H_0: mean(\text{PCIT-T}) = mean(\text{CoS-P}) = mean(\text{Waitlist})$ -->
<!-- $H_A: mean(\text{PCIT-T}) > mean(\text{CoS-P}) > mean(\text{Waitlist})$ -->

<!-- This is a joint hypothesis test. -->

<!-- First, confirming that the ordering of the means is as hypothesized: -->

<!-- ```{r, echo=FALSE} -->
<!-- result.change.d -->
<!-- ``` -->

<!-- So, the mean change for PCITT > COSP > Waitlist. -->

<!-- Now, to get the p-value associated with the joint hypothesis. -->


<!-- ```{r, echo=FALSE, warning=FALSE} -->

<!-- ## Specify contrast matrices -->
<!-- # Hyp1 is 1 df test -->
<!-- # could use alternative="less" ir alternative="greater" in glht -->
<!-- hyp1   <- matrix(c(0, 0, -1,0, 0,0, -1,0,0 ),1) -->
<!-- # Hyp2 is 1 df test,  -->
<!-- hyp2   <- matrix(c(0, 0,  1,-1, 0,0, 1,-1,0 ),1) -->

<!-- # hypF is combined 2df test. Yes! -->
<!-- h<-rbind(hyp1,hyp2) -->
<!-- hypF <- as.matrix(h) -->
<!-- GlobalF <- glht(result3, linfct = hypF, test = Ftest()) -->
<!-- GlobalF$linfct -->
<!-- summary(GlobalF, test = Ftest()) -->

<!-- #also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html -->

<!-- ``` -->

<!-- p>0.05, so we cannot reject the null hypothesis. -->


<!-- #### Hypothesis 2 -->

<!-- There will be a gradient of size of change from T2 to T3 such that the change in the PCIT-T group > the change in the CoS-P group.  This isn't really a gradient; it's just 1 hypothesis. -->

<!-- First, the changw in PCITT and COSP from time 2 to time 3: -->

<!-- ```{r, echo=FALSE} -->
<!-- result.change.d -->
<!-- ``` -->

<!-- There is greater change in PCITT( -1.16) than in COSP (1.87) -->


<!-- ```{r, echo=FALSE, warning=FALSE} -->

<!-- ## Specify contrast matrices -->
<!-- # Hyp1 is 1 df test -->
<!-- # could use alternative="less" ir alternative="greater" in glht -->
<!-- hyp2   <- matrix(c(0, 0,  0,0, 0,0, -1,0,1 ),1) -->
<!-- hyp2F <- glht(result3, linfct = hyp2) -->
<!-- summary(hyp2F) -->
<!-- #also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html -->

<!-- ``` -->

<!-- p>0.05, so we do not have enough evidence to suggest that the change from t2 to t3 in PCITT is greater than the change in COSP. -->





#### Residuals

QQ plot does show some deviation from normality. You can see in the histogram that the tails are long on both ends, but there is symmetry.

```{r, echo=FALSE, warning=FALSE}
plot(result5)

qqnorm(residuals(result5))
qqline(residuals(result5), col="red")

hist((residuals(result5)))


# plot residuals against predictors

ggplot(data.frame(x1=dat$group,pearson=residuals(result5,type="pearson")),
       aes(x=x1,y=pearson)) +
  geom_point() +
  theme_bw()

ggplot(data.frame(x2=dat$time,pearson=residuals(result5,type="pearson")),
       aes(x=x2,y=pearson)) +
  geom_point() +
  theme_bw()

ggplot(data.frame(x2=dat$lie,pearson=residuals(result5,type="pearson")),
       aes(x=x2,y=pearson)) +
  geom_point() +
  theme_bw()

```

```{r}
rm(list = ls())
```
