---
title: "PCIT-CoS-Waitlist RCT Paper 1: Parent & Child Behaviour Outcomes"
author: "Nancy Briggs"
date: "05 JAN 2021"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(tidyselect)
library(SASmarkdown)
library(nlme)
library(emmeans)
library(multcomp)
library(kableExtra)
library(ggplot2)



options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)

```

# Intro

This document contains the results for the analyses described in [Kohlhoff et al (2020)](https://doi.org/10.1186/s40359-020-00457-7). Outcomes are for Paper 1, Parent & Child Behaviour Outcomes.

Analyses were done in R. 

R Core Team (2021). R: A language and environment for statistical computing. R Foundation for Statistical Computing, Vienna, Austria. URL https://www.R-project.org/.


## Description of the RCT

This study is a block-randomized, longitudinal RCT to evaluate the efficacy of Parent Child Interaction Therapy (PCIT-T) for children aged 14-24 months with disruptive behaviors. Ninety children were randomized to one of three treatment arms: PCIT-T, Circle of Security-Parenting (CoS-P) or Waitlist Control. Outcomes of interest arem

- Parents
  - Observed Positive Parenting (DPICS Labelled Praise, % of total parent verbalisations)
  - Observed Negative Parenting (DPICS Negative Tale, % of total parent verbalisations)
  - Child abuse potential (BCAP)

- Children
   - Child Externalising Behaviour (CBCL)
   - Child Internalising Behaviour (CBCL)
   - Child Temperament (ECBQ Fear)
   - Child Temperament (ECBQ Shyness)
   - Child Social-Emotional Functioning (BITSEA)
   - Child Social-Emotional Functioning (DECA)
   - Child Compliance (DPICS Child Compliance, % of total parent effective commands)



## Study design

Families referred to the Karitane Toddle Clinic (KTC) with children aged 14-24 months were randomised toone of 3 treatment arms: PCIT-T, CS-P or Waitlist.  All participants completed Baseline (T1) measures.  PCIT-T and CoS-P groups participated in the interventions and them provided post-treatment (T2) assessments 8 weeks after T1. For the Waitlist group, their T2 measurement occurred 8 weeks after T1, but with no intervention (post-waitlist). T3 assessments were measured 4 months after the end of treatment.

## Hypotheses

We thus hypothesize that from baseline (‘T1’) to
post-treatment/waitlist (‘T2’) there will be a gradient
effect, whereby parent-child dyads who receive PCIT-T
will show the greatest gains and parent-child dyads who
receive no treatment will show the least gains with respect
to the following variables:

  * Improved parenting capacity through increased 
    + increased labeled praise of the child
    + increased parenting sensitivity
    + increased parent sense of competence in managing negative toddler emotions
    + increased parent mentalizing about the child
    + increased parent emotion regulation
    + decreased negative statements directed toward the child
    + decreased parent sense of caregiving helplessness
    + decreased child abuse potential
    + decreased parental stress
    
  * Improved child outcomes through
    + increased emotion regulation maturity
    + increased social-emotional functioning including initiative, relationship functioning and self-regulation
    + increased attachment security
    + decreased externalizing behavior


Second, we expect that for the two active treatment conditions, this gradient effect will be maintained at a 4-month post-treatment follow-up, whereby parent-child dyads that receive PCIT-T will show superior outcomes on all variables compared to parent-child dyads where the parent received COS-P.


## General Statistical Approach

Analyses will be done on an intention-to-treat principle. Results with p-values < 0.05 will be considered significant -- Multiple Comparisons?  Estimates of treatment effects will be presented with 95% confidence intervals.

In general, a linear mixed model will be used to analyse the data at T2 & T3. A random effect of individual will be specified, and fixed effects of measurement time and treatment group and their interaction will be included.

Twisk J, Bosman L, Hoekstra T, Rijnhart J, Welten M, Heymans M. (2018) Different ways to estimate treatment effects in randomised controlled trials. Contemporary Clinical Trials Communications, 10, 80-85, https://doi.org/10.1016/j.conctc.2018.03.008.



Stated hypotheses are:

1. There will be a gradient of size of change from T1 to T2 such that the mean in the PCIT-T group > the mean in the CoS-P group > the mean in the Waitlist group.

Estimates of the change from T1 to T2 within each group will be obtained along with 95% CI.  In addition, an overall test of the difference in change between PCIT-T and CoS-P and between CoS-P and Waitlist will be performed. 
<!-- (CONTRAST statement in SAS, including both ESTIMATE codes.) -->


2. There will be a gradient of size of change from T2 to T3 such that the change in the PCIT-T group > the change in the CoS-P group.

Estimates of the change from T2 to T3 within each group will be obtained along with 95% CI.  In addition, an overall test of the difference in change between PCIT-T and CoS-P and between CoS-P and Waitlist will be performed. 
<!-- (CONTRAST statement in SAS, including both ESTIMATE codes.) -->

```{r, echo=FALSE}
# Setting up output tables for summary at end.
# Table 1 demographics total and within group
# Descriptives for outcomes by group
# ANOVA Unadjusted
# EMM Unadjusted
# ANOVA Adjusted for ASDcase
# EMM Adjusted for ASDcase
# (?) ANOVA Adjusting for ASDcase & Total Therapy Time
# (?) EMM Adjusting for ASDcase & Total Therapy Time

```

# Analyses Unadjusted

<!-- Note: add in this if needed to get pvalues for coefficients -->

<!-- library(lmerTest) -->
<!-- m.test <- as(result7,"merModLmerTest") -->
<!-- print(summary(m.test,ddf="lme4"),correlation=FALSE) -->
<!-- print(summary(m.test,ddf="Satterthwaite"),correlation=FALSE) -->

<!-- ![CBCL Externalizing](CBCLExt/analysis.html) -->
<!-- ![CBCL Internalizing](CBCLInt/analysis.html) -->

![BCAP](BCAP/analysis.html)

<!-- ```{r, child="CBCLExt/analysis.Rmd"} -->
<!-- #DONE 25Jan2022 -->
<!-- ``` -->

<!-- ```{r, child="CBCLInt/analysis.Rmd"} -->
<!-- #DONE 25Jan2022 -->
<!-- ``` -->

<!-- ```{r, child="BCAP/analysis.rmd"} -->
<!-- ``` -->

<!-- ```{r, child="BCAP/analysis_lie_TV.rmd"} -->
<!-- #DONE 31Jan2022 -->
<!-- ``` -->

<!-- ```{r, child="BCAP/analysis_baselineLie.rmd"} -->
<!-- #DONE 25Jan2022 -->
<!-- ``` -->

<!-- ```{r, child="DPICSlp/analysis.rmd"} -->
<!-- #DONE 31Jan2022 -->
<!-- ``` -->

<!-- ```{r, child="DPICSnta/analysis.rmd"} -->
<!-- #DONE 31Jan2022 -->
<!-- ``` -->


<!-- ```{r, child="DPICSco/analysis.rmd"} -->
<!-- #DONE 01Feb2022 -->
<!-- ``` -->


# Analyses Adjusted for ASD

<!-- ```{r, child="CBCLExt/analysisASD.Rmd"} -->
<!-- #DONE 25Jan2022 -->
<!-- ``` -->

<!-- ```{r, child="CBCLInt/analysisASD.Rmd"} -->
<!-- #DONE 25Jan2022 -->
<!-- ``` -->

<!-- ```{r, child="DPICSlp/analysis_asd.rmd"} -->
<!-- #DONE 31Jan2022 -->
<!-- ``` -->

<!-- ```{r, child="DPICSnta/analysis_asd.rmd"} -->
<!-- #DONE 31Jan2022 -->
<!-- ``` -->

<!-- ```{r, child="DPICSco/analysis_ASD.rmd"} -->
<!-- #DONE 01Feb2022 -->
<!-- ``` -->

