---
title: "PCIT-CoS-Waitlist RCT Paper 1:  Total Child Distress "
author: "Nancy Briggs"
date: "05 JAN 2021"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---


### Total Child Distress {.tabset .tabset-fade .tabset-pills}

Pre_Total_Distress
Post_Total_Distress
Fup_Total_Distress
 
 


```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)

library(tidyverse)
library(tidyselect)
library(SASmarkdown)
library(nlme)
library(emmeans)
library(multcomp)
library(kableExtra)
library(ggplot2)
library(lmerTest)
library(effectsize)
library(boot)
library(haven)
options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)

```


```{r, echo=FALSE}
# Data 
# Import data
# recoding for CBCL and totals done in SPSS before bringing data in to R

dataer <- read_sav("../PCITT RCT main dataset 2.4.24 - ER data only.sav")
names(dataer) <- tolower(names(dataer))
data <- read.csv("../data_prep.csv", fileEncoding="UTF-8-BOM")
names(data) <- tolower(names(data))

data<- data %>%
  full_join(.,dataer, by=c("id", "group"))

# keep relevant demographic and outcome variables

data <- data %>%
  dplyr::select(id:treatment_time_mins, dobchild, sexchild, numbhousehold, incfamily,

                
                
                pre_total_distress, post_total_distress, fup_total_distress
 
                )


```




#### Descriptives

```{r, echo=FALSE}
#    pre_total_distress, post_total_distress, fup_total_distress

desctable <- data %>%
  group_by(group) %>%
  mutate(score1 = case_when(
    pre_total_distress == -9 ~ NA_real_,          #assign missing values to zeros
    TRUE ~ pre_total_distress)) %>%
  mutate(score2 = case_when(
    post_total_distress == -9 ~ NA_real_,         #assign missing values to zeros
    TRUE ~ post_total_distress)) %>%
  mutate(score3 = case_when(
    fup_total_distress == -9 ~ NA_real_,           #assign missing values to zeros
    TRUE ~ fup_total_distress)) %>%
  summarise(n1 = sum(!is.na(score1)) , 
            mean1=mean(pre_total_distress, na.rm = T ),
            sd1=sd(pre_total_distress, na.rm = T ),
            median1=median(pre_total_distress, na.rm = T ),
            min1=min(pre_total_distress, na.rm = T ),
            max1=max(pre_total_distress, na.rm = T ),
            
            n2 = sum(!is.na(score2)) , 
            mean2=mean(post_total_distress, na.rm = T ),
            sd2=sd(post_total_distress, na.rm = T ),
            median2=median(post_total_distress, na.rm = T ),
            min2=min(post_total_distress, na.rm = T ),
            max2=max(post_total_distress, na.rm = T ),
            
            n3 = sum(!is.na(score3)) , 
            mean3=mean(fup_total_distress, na.rm = T ),
            sd3=sd(fup_total_distress, na.rm = T ),
            median3=median(fup_total_distress, na.rm = T )
            # ,
            # min3=min(fu_er_emotional_maturity, na.rm = T ),
            # max3=max(fu_er_emotional_maturity, na.rm = T )  
            ) %>%
  pivot_longer(cols = -group,  
                names_to = c('.value', 'time'), 
                    names_pattern = '(.*)(\\d+)')

write.csv(desctable,"desctable.csv", row.names = TRUE)

```

```{r, echo=FALSE}
desctable %>%
           kbl(digits=3) %>%
           kable_minimal()
```



```{r, echo=FALSE}

# Histograms
# plot
p <- data %>%
  dplyr::select(id,group,pre_total_distress, post_total_distress, fup_total_distress
) %>%
  pivot_longer( cols = c("pre_total_distress", 
                "post_total_distress", "fup_total_distress"), 
                names_to="t", values_to="outcome") %>%
  mutate(time = ifelse(t == "pre_total_distress",  1,
                ifelse(t == "post_total_distress", 2, 3))) %>%
  mutate(log=log(outcome+.5))
  

  
hist<-ggplot(p, aes(x=outcome) )+
  geom_histogram(alpha=0.6, binwidth = 5) +
  xlab("Total Child Distress") +
  ylab("Frequency") +
  facet_wrap(group~time)

group.labs <- c("PCIT-T", "COS-P", "Waitlist")
names(group.labs) <- c("1", "2", "3")

# New facet label names for group variable
time.labs <- c("Pre", "Post","FollowUp")
names(time.labs) <- c("1", "2","3")
# Create the plot
hist + theme_minimal() +
  facet_grid(
  group ~ time, 
  labeller = labeller(group = group.labs, time = time.labs) )

histlog<-ggplot(p, aes(x=log) )+
  geom_histogram(alpha=0.6, binwidth = 1) +
  xlab("Total Child Distress") +
  ylab("Frequency") +
  facet_wrap(group~time)
# Create the plot
histlog + theme_minimal() +
  facet_grid(
  group ~ time, 
  labeller = labeller(group = group.labs, time = time.labs) )

```

#### Model & Results

A linear mixed model with random effect for individual and fixed effects of group, time and the interaction was run. Estimated marginal means were obtained, and contrasts specified to test Hypothesis 1 and Hypothesis 2.

Log transformed data werwe used due to violations of normality and homogeneity of variance of residuals when using the raw outcome variable.

```{r, echo=FALSE, warning=FALSE}
#  analysis
dat <- data %>%
  dplyr::select(id,group,pre_total_distress, post_total_distress, fup_total_distress
) %>%
  pivot_longer( cols = c("pre_total_distress", 
                "post_total_distress", "fup_total_distress"), 
                names_to="t", values_to="outcome") %>%
  mutate(time = ifelse(t == "pre_total_distress",  1,
                ifelse(t == "post_total_distress", 2, 3))) %>%
  mutate(log=log(outcome+.5))

dat$time<- as.factor(dat$time)
dat$group<- as.factor(dat$group)

# result0<-lmer(outcome ~ (1 | id), dat)
# result1<-lmer(outcome ~ group + (1 | id), dat)
# result2<-lmer(outcome ~ group + time + (1 | id), dat)
# result3<-lmer(outcome ~ group*time + (1 | id), dat)

result0<-lmer(log ~ (1 | id), dat)
result1<-lmer(log ~ group + (1 | id), dat)
result2<-lmer(log ~ group + time + (1 | id), dat)
result3<-lmer(log ~ group*time + (1 | id), dat)

```




This is the ANOVA table for the model. p-values are likelihood ratio tests for adding the new term to the model just before. A p-value <0.05 indicates that adding that term results in better model fit. The most important p-value is the interaction p-value. Here, there is no evidence that the pattern of change over time is different between the groups (p=0.125).

```{r, echo=FALSE}
anovatab<-anova(result0,result1,result2,result3)

anovatab <-anovatab %>%
           mutate(Effect = ifelse(npar == 3, "Intercept",
                           ifelse(npar == 5, "Group",
                           ifelse(npar == 7, "Time",
                           ifelse(npar == 10, "Group*Time",""))))) %>%
           relocate(Effect) 
anovatab %>%
           kbl(digits=3) %>%
           kable_styling()
```



```{r}
summary(result3)
```



#### Estimated Means & Change

Estimated means for each Treatment and Time are presented. In addition, within-group
change between time points is presented with 95% CIs.


```{r child='bootstrapped_es_CI.Rmd'}
```


#### Hypothesis 1

We hypothesize that from baseline (‘T1’) to post-treatment/waitlist (‘T2’) there will be a gradient effect, whereby parent-child dyads who receive PCIT-T
will show the greatest gains and parent-child dyads who
receive no treatment will show the least gains. For the outcome, then, the 
null hypothesis is:

$H_0:  \text {mean change}(\text{PCIT-T}) = \text {mean change}(\text{CoS-P}) = \text {mean change}(\text{Waitlist})$
$H_A: \text {mean change}(\text{PCIT-T}) > \text {mean change}(\text{CoS-P}) > \text {mean change}(\text{Waitlist})$

This is a joint hypothesis test.

First, confirming that the ordering of the means is as hypothesized:

```{r, echo=FALSE}
result.change.d
```

Now, to get the p-value associated with the joint hypothesis.

```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp1   <- matrix(c(0, 0,0, 0,0, -1,0,0 ),1)
# Hyp2 is 1 df test, 
hyp2   <- matrix(c(0, 0,0, 0,1, 0,1,0 ),1)

# hypF is combined 2df test. Yes!
h<-rbind(hyp1,hyp2)
hypF <- as.matrix(h)
GlobalF <- glht(result3, linfct = hypF, test = Ftest())
GlobalF$linfct
summary(GlobalF, test = Ftest())

#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html

```
p>.05



#### Hypothesis 2
##### Change from T2-T3

There will be a gradient of size of change from T2 to T3 such that the change in the PCIT-T group > the change in the CoS-P group.  This isn't really a gradient; it's just 1 hypothesis.

First, the change in PCITT and COSP from time 2 to time 3:

```{r, echo=FALSE}
result.change.d
```



```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, -1,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2)
summary(hyp2F)
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html

```

p>0.05 so we do not have evidence that the change from t2-t3 is different between PCITT and COSP

##### Change from T1-T3


First, the change in PCITT and COSP from time 1 to time 3:

```{r, echo=FALSE}
result.change.d
```



```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, 0,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2)
summary(hyp2F)
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html

```

p>0.05 so we do not have evidence that the change from t1-t3 is different between PCITT anf COSP

#### Residuals

Assumptions of the model are met.

```{r, echo=FALSE, warning=FALSE}
plot(result3)

qqnorm(residuals(result3))
qqline(residuals(result3), col="red")

# plot residuals against predictors
dat <- na.omit(dat)
ggplot(data.frame(x1=dat$group,pearson=residuals(result3,type="pearson")),
       aes(x=x1,y=pearson)) +
  geom_point() +
  theme_bw()

ggplot(data.frame(x2=dat$time,pearson=residuals(result3,type="pearson")),
       aes(x=x2,y=pearson)) +
  geom_point() +
  theme_bw()


```

```{r}
rm(list = ls())
```
