---
title: "PCIT-CoS-Waitlist RCT Paper 1: Controlled Verbosity Attuned"
author: "Nancy Briggs"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---


### Controlled Verbosity Attuned {.tabset .tabset-fade .tabset-pills}

Output run on `r Sys.Date()`



Variables: Attuned_controlled_verbosity Attuned_controlled_verbosity_t2 Attuned_controlled_verbosity_t3

But these are percentages, created from 
Attuned_controlled_verbosity = `Attuned` / `MM_total_codes`


```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=1)
library(lme4)

library(tidyverse)
library(tidyselect)
library(SASmarkdown)
library(nlme)
library(emmeans)
library(multcomp)
library(kableExtra)
library(ggplot2)
library(lmerTest)
library(effectsize)
library(boot)
library(DHARMa)

options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)

```


```{r, echo=FALSE}
# Data 
# Import data
# recoding for CBCL and totals done in SPSS before bringing data in to R

data <- read.csv("../data_prep.csv", fileEncoding="UTF-8-BOM")
names(data) <- tolower(names(data))

# keep relevant demographic and outcome variables

data <- data %>%
  dplyr::select(id:treatment_time_mins, dobchild, sexchild, numbhousehold, incfamily,


                #Controlled Verbosity Attuned
                attuned_controlled_verbosity,
                attuned_controlled_verbosity_t2,
                attuned_controlled_verbosity_t3,
                
                # Attuned
                mm_attuned, t2_attuned, t3_attuned,
                # MM totals  
                mm_total_codes, t2_mm_total_codes,t3_mm_total_codes
                )


```




#### Descriptives

```{r, echo=FALSE}
#   attuned_controlled_verbosity,
#                attuned_controlled_verbosity_t2,
#                attuned_controlled_verbosity_t3

desctable <- data %>%
  group_by(group) %>%
  summarise(n1 = sum(!is.na(attuned_controlled_verbosity)) , 
            mean1=mean(attuned_controlled_verbosity, na.rm = T ),
            sd1=sd(attuned_controlled_verbosity, na.rm = T ),
            median1=median(attuned_controlled_verbosity, na.rm = T ),
            min1=min(attuned_controlled_verbosity, na.rm = T ),
            max1=max(attuned_controlled_verbosity, na.rm = T ),
            
            n2 = sum(!is.na(attuned_controlled_verbosity_t2)) , 
            mean2=mean(attuned_controlled_verbosity_t2, na.rm = T ),
            sd2=sd(attuned_controlled_verbosity_t2, na.rm = T ),
            median2=median(attuned_controlled_verbosity_t2, na.rm = T ),
            min2=min(attuned_controlled_verbosity_t2, na.rm = T ),
            max2=max(attuned_controlled_verbosity_t2, na.rm = T ),
            
            n3 = sum(!is.na(attuned_controlled_verbosity_t3)) , 
            mean3=mean(attuned_controlled_verbosity_t3, na.rm = T ),
            sd3=sd(attuned_controlled_verbosity_t3, na.rm = T ),
            median3=median(attuned_controlled_verbosity_t3, na.rm = T ),
            min3=min(attuned_controlled_verbosity_t3, na.rm = T ),
            max3=max(attuned_controlled_verbosity_t3, na.rm = T )  ) %>%
  pivot_longer(cols = -group,  
                names_to = c('.value', 'time'), 
                    names_pattern = '(.*)(\\d+)')

write.csv(desctable,"desctable.csv", row.names = TRUE)

```

```{r, echo=FALSE}
desctable %>%
           kbl(digits=3) %>%
           kable_minimal()
```

Removing the G3 T3 observation


```{r, echo=FALSE, warning=FALSE}
#   attuned_controlled_verbosity,
#                attuned_controlled_verbosity_t2,
#                attuned_controlled_verbosity_t3

data <- data %>%
  mutate(attuned_controlled_verbosity_t3=ifelse(group==3,NA,attuned_controlled_verbosity_t3))

desctable <- data %>%
  group_by(group) %>%
  summarise(n1 = sum(!is.na(attuned_controlled_verbosity)) , 
            mean1=mean(attuned_controlled_verbosity, na.rm = T ),
            sd1=sd(attuned_controlled_verbosity, na.rm = T ),
            median1=median(attuned_controlled_verbosity, na.rm = T ),
            min1=min(attuned_controlled_verbosity, na.rm = T ),
            max1=max(attuned_controlled_verbosity, na.rm = T ),
            
            n2 = sum(!is.na(attuned_controlled_verbosity_t2)) , 
            mean2=mean(attuned_controlled_verbosity_t2, na.rm = T ),
            sd2=sd(attuned_controlled_verbosity_t2, na.rm = T ),
            median2=median(attuned_controlled_verbosity_t2, na.rm = T ),
            min2=min(attuned_controlled_verbosity_t2, na.rm = T ),
            max2=max(attuned_controlled_verbosity_t2, na.rm = T ),
            
            n3 = sum(!is.na(attuned_controlled_verbosity_t3)) , 
            mean3=mean(attuned_controlled_verbosity_t3, na.rm = T ),
            sd3=sd(attuned_controlled_verbosity_t3, na.rm = T ),
            median3=median(attuned_controlled_verbosity_t3, na.rm = T ),
            min3=min(attuned_controlled_verbosity_t3, na.rm = T ),
            max3=max(attuned_controlled_verbosity_t3, na.rm = T )  ) %>%
  pivot_longer(cols = -group,  
                names_to = c('.value', 'time'), 
                    names_pattern = '(.*)(\\d+)')

write.csv(desctable,"desctable_exclG3.csv", row.names = TRUE)

```

```{r, echo=FALSE}
desctable %>%
           kbl(digits=3) %>%
           kable_minimal()
```



```{r, echo=FALSE}
##   attuned_controlled_verbosity, attuned_controlled_verbosity_t2, attuned_controlled_verbosity_t3
# Histograms
# plot
p <- data %>%
  dplyr::select(id,group, attuned_controlled_verbosity, attuned_controlled_verbosity_t2, attuned_controlled_verbosity_t3) %>%
  pivot_longer( cols = c("attuned_controlled_verbosity", "attuned_controlled_verbosity_t2", "attuned_controlled_verbosity_t3"),  names_prefix = "ders_", 
                names_to="t", values_to="outcome") %>%
  mutate(time = ifelse(t == "attuned_controlled_verbosity",  1,
                ifelse(t == "attuned_controlled_verbosity_t2", 2, 3))) 
  

  
hist<-ggplot(p, aes(x=outcome) )+
  geom_histogram(alpha=0.6, binwidth = .01) +
  xlab("Controlled Verbosity Attuned") +
  ylab("Frequency") +
  facet_wrap(group~time)

group.labs <- c("PCIT-T", "COS-P", "Waitlist")
names(group.labs) <- c("1", "2", "3")

# New facet label names for group variable
time.labs <- c("Pre", "Post","FollowUp")
names(time.labs) <- c("1", "2","3")
# Create the plot
hist + theme_minimal() +
  facet_grid(
  group ~ time, 
  labeller = labeller(group = group.labs, time = time.labs) )

```

#### Model & Results

Controlled verbosity is a ratio of two values: Attuned and MM Total.  As such a binomial GLMM will be used.

A generalized linear mixed model with binomial distribution and  random effect for individual and fixed effects of group, time and the interaction was run. Estimated marginal means were obtained, and contrasts specified to test Hypothesis 1 and Hypothesis 2.

```{r, echo=FALSE, warning=FALSE}
#  analysis
data <- data %>%
  mutate(attuned_controlled_verbosity_t3=ifelse(group==3,NA,attuned_controlled_verbosity_t3)) %>%
    mutate(t3_mm_total_codes=ifelse(group==3,NA,t3_mm_total_codes))

dat <- data %>%
  dplyr::select(id,group,attuned_controlled_verbosity, attuned_controlled_verbosity_t2,
                attuned_controlled_verbosity_t3) %>%
  pivot_longer( cols = c("attuned_controlled_verbosity", "attuned_controlled_verbosity_t2", "attuned_controlled_verbosity_t3"),  names_prefix = "bitseaprob_", 
                names_to="t", values_to="acv") %>%
  mutate(time = ifelse(t == "attuned_controlled_verbosity",  1,
                ifelse(t == "attuned_controlled_verbosity_t2", 2, 3)))
dat1 <- data %>%
  dplyr::select(id,group,mm_attuned, t2_attuned, t3_attuned) %>%
  pivot_longer( cols = c("mm_attuned", "t2_attuned", "t3_attuned"),  names_prefix = "bitseaprob_", 
                names_to="t", values_to="attuned") %>%
  mutate(time = ifelse(t == "mm_attuned",  1,
                ifelse(t == "t2_attuned", 2, 3)))
dat2 <- data %>%
  dplyr::select(id,group,mm_total_codes, t2_mm_total_codes,t3_mm_total_codes) %>%
  pivot_longer( cols = c("mm_total_codes", "t2_mm_total_codes","t3_mm_total_codes"), 
                names_prefix = "bitseaprob_", 
                names_to="t", values_to="mm") %>%
  mutate(time = ifelse(t == "mm_total_codes",  1,
                ifelse(t == "t2_mm_total_codes", 2, 3)))
dat <- dat %>%
  left_join(.,dat1, by=c("id","time")) %>%
  left_join(.,dat2, by=c("id","time"))%>%
  filter(!is.na(acv) )


dat$time<- as.factor(dat$time)
dat$group<- as.factor(dat$group)

# transform 0 values to (y * (n−1) + 0.5) / n
dat <- dat %>%
  mutate(acv=ifelse(acv==0, .5 / 88, acv)) 
#  mutate(acv=ifelse(attuned<=1, .05, acv)) 


result0<-glmer( acv ~ (1 | id), family="binomial",weight=mm, dat) 
result1<-glmer( acv ~ group + (1 | id),family="binomial",weight=mm, dat) 
result2<-glmer( acv ~ group + time + (1 | id), family="binomial",weight=mm, dat) 
result3<-glmer( acv ~ group*time + (1 | id), family="binomial",weight=mm, dat) 

```





```{r, echo=FALSE}
anovatab<-anova(result0,result1,result2,result3)

anovatab <-anovatab %>%
           mutate(Effect = ifelse(npar == 2, "Intercept",
                           ifelse(npar == 4, "Group",
                           ifelse(npar == 6, "Time",
                           ifelse(npar == 9, "Group*Time",""))))) %>%
           relocate(Effect) 
anovatab %>%
           kbl(digits=3) %>%
           kable_styling()


## Extract pvaue for evidence / no evidence
evidence = ifelse(anovatab[4,9]<0.05, "evidence", "no evidence")
```



This is the ANOVA table for the model. p-values are likelihood ratio tests for adding the new term to the model just before. A p-value <0.05 indicates that adding that term results in better model fit. The most important p-value is the interaction p-value. Here, there is `r evidence` that the pattern of change over time is different among the groups (p=`r round(anovatab[4,9],3)`).




```{r}
summary(result3)
```



#### Estimated Means & Change

Estimated means for each Treatment and Time are presented. In addition, within-group
change between time points is presented with 95% CIs.

Note that the "effect size" is the OR.  Interpret the probabilities, and the tests of difference but note really the OR.


```{r child='bootstrapped_es_CI.Rmd'}
```

```{r child='TestingHypotheses.Rmd'}
```



#### Residuals

Assumptions of the model are met.

```{r, echo=FALSE, warning=FALSE}
sim= simulateResiduals(result3)
resid = residuals(sim, quantileFunction = qnorm)


# This line may have to be:
plot(resid~predict(result3, re.form = NA,type="response", outliers=c(-7.7)))


plot(result3,form = resid(., type = "pearson") ~ predict(., re.form = NA) )


qqnorm(residuals(result3))
qqline(residuals(result3), col="red")

# plot residuals against predictors
dat <- na.omit(dat)
ggplot(data.frame(x1=dat$group,pearson=residuals(result3,type="pearson")),
       aes(x=x1,y=pearson)) +
  geom_point() +
  theme_bw()

ggplot(data.frame(x2=dat$time,pearson=residuals(result3,type="pearson")),
       aes(x=x2,y=pearson)) +
  geom_point() +
  theme_bw()


```

```{r}
rm(list = ls())
```