---
title: "PCIT-CoS-Waitlist RCT Paper 1: DPICS LP"
author: "Nancy Briggs"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---


### DPICS LP  {.tabset .tabset-fade .tabset-pills}

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(lme4)
library(tidyverse)
library(tidyselect)
library(SASmarkdown)
library(nlme)
library(emmeans)
library(multcomp)
library(kableExtra)
library(ggplot2)
library(lmerTest)


options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)

```


```{r, echo=FALSE}
# Data 
# Import data
# recoding for CBCL and totals done in SPSS before bringing data in to R

data <- read.csv("../data_prep.csv", fileEncoding="UTF-8-BOM")
names(data) <- tolower(names(data))

# keep relevant demographic and outcome variables

data <- data %>%
  dplyr::select(id:treatment_time_mins, dobchild, sexchild, numbhousehold, incfamily,
                #DPICS Labelled Praise
                parent_lp_perc_total_time_1, parent_lp_perc_total_time_2, parent_lp_perc_total_time_3,
                #DPICS Negative Talk
                parent_nta_perc_total_time_1, parent_nta_perc_total_time_2, parent_nta_perc_total_time_3,
               
                
                #Child Compliance
                total_co_perc_time_1,total_co_perc_time_2,total_co_perc_time_3
                
                # others, but not essential for paper
                #3. Child Temperament (ECBQ)
                #4. Child social-emotional functioning - #1: Brief Infant Toddler Social Emotional Assessment (BITSEA)
                #5. Child social-emotional functioning - #2: Child Devereux Early Childhood Assessment (DECA)
                
                )


### import total time
time <- read.csv("../data_prep_DPiCStime.csv", fileEncoding="UTF-8-BOM")
names(time) <- tolower(names(time))
time <- time %>%
  dplyr::select(id, parent_total_verbalisations_time_1, parent_total_lp_time_1, 
                  parent_total_nta_time_1, total_eff_commands_time_1, 
                  child_co_total_time_1, child_nc_total_time_1, 
                  total_co_perc_time_1, total_nc_perc_time_1, 
                  parent_total_verbalisations_time_2, parent_total_lp_time_2, 
                  parent_total_nta_time_2, total_eff_commands_time_2, 
                  child_co_total_time_2, child_nc_total_time_2, 
                  total_co_perc_time_2, total_nc_perc_time_2, 
                  parent_total_verbalisations_time_3, parent_total_lp_time_3, 
                  parent_total_nta_time_3, total_eff_commands_time_3, 
                  child_co_total_time_3, child_nc_total_time_3, 
                  total_co_perc_time_3, total_nc_perc_time_3	)

# Join
data <- left_join(data, time, by="id")


```




#### Descriptives

```{r, echo=FALSE, warning=F}
# parent_lp_perc_total_time_1, parent_lp_perc_total_time_2, parent_lp_perc_total_time_3,


desctable <- data %>%
  group_by(group) %>%
  mutate(score1 = case_when( 
    parent_lp_perc_total_time_1 == -9 ~ NA_real_,                                #assign missing values to zeros
    TRUE ~ parent_lp_perc_total_time_1)) %>%
  mutate(score2 = case_when( 
    parent_lp_perc_total_time_2 == -9 ~ NA_real_,                                #assign missing values to zeros
    TRUE ~ parent_lp_perc_total_time_2)) %>%
  mutate(score3 = case_when( 
    parent_lp_perc_total_time_3 == -9 ~ NA_real_,                                #assign missing values to zeros
    TRUE ~ parent_lp_perc_total_time_3)) %>%
  summarise(n1 = sum(!is.na(score1)) , 
            mean1=mean(parent_lp_perc_total_time_1, na.rm = T ),
            sd1=sd(parent_lp_perc_total_time_1, na.rm = T ),
            median1=median(parent_lp_perc_total_time_1, na.rm = T ),
            min1=min(parent_lp_perc_total_time_1, na.rm = T ),
            max1=max(parent_lp_perc_total_time_1, na.rm = T ),
            
            n2 = sum(!is.na(score2)) , 
            mean2=mean(parent_lp_perc_total_time_2, na.rm = T ),
            sd2=sd(parent_lp_perc_total_time_2, na.rm = T ),
            median2=median(parent_lp_perc_total_time_2, na.rm = T ),
            min2=min(parent_lp_perc_total_time_2, na.rm = T ),
            max2=max(parent_lp_perc_total_time_2, na.rm = T ),
            
            n3 = sum(!is.na(score3)) , 
            mean3=mean(parent_lp_perc_total_time_3, na.rm = T ),
            sd3=sd(parent_lp_perc_total_time_3, na.rm = T ),
            median3=median(parent_lp_perc_total_time_3, na.rm = T ),
            min3=min(parent_lp_perc_total_time_3, na.rm = T ),
            max3=max(parent_lp_perc_total_time_3, na.rm = T )  ) %>%
  pivot_longer(cols = -group,  
                names_to = c('.value', 'time'), 
                    names_pattern = '(.*)(\\d+)')

write.csv(desctable,"desctable.csv", row.names = TRUE)




```




_Descriptives of Percent of Total Verbalisations_

```{r, echo=FALSE}
desctable %>%
           kbl() %>%
           kable_styling()
```



```{r, echo=FALSE}

# Histograms
# plotv parent_lp_perc_total_time_1
p <- data %>%
  dplyr::select(id,group, parent_lp_perc_total_time_1, parent_lp_perc_total_time_2, parent_lp_perc_total_time_3) %>%
  pivot_longer( cols = starts_with("parent_lp_perc_total_time_"), names_prefix = "time", 
                names_to="t", values_to="outcome") %>%
  mutate(time = ifelse(t == "parent_lp_perc_total_time_1",  1,
                ifelse(t == "parent_lp_perc_total_time_2", 2, 3))) 
  

  
hist<-ggplot(p, aes(x=outcome) )+
  geom_histogram(alpha=0.6, binwidth = 5) +
  xlab("DPICS Percent LP") +
  ylab("Frequency") +
  facet_wrap(group~time)

group.labs <- c("PCIT-T", "COS-P", "Waitlist")
names(group.labs) <- c("1", "2", "3")

# New facet label names for group variable
time.labs <- c("Pre", "Post","FollowUp")
names(time.labs) <- c("1", "2","3")
# Create the plot
hist + theme_minimal() +
  facet_grid(
  group ~ time, 
  labeller = labeller(group = group.labs, time = time.labs) )

```


#### Model & Results

A generalized linear mixed model (GLMM) with random effect for individual and fixed effects of group, time and the interaction was run on the proportion of time with LP responses. A binomial distribution was specified with a logit link. Estimated marginal means were obtained, and contrasts specified to test Hypothesis 1 and Hypothesis 2. Assumptions on the residuals were examined using simulated residuals (see below).


```{r, echo=FALSE}
#  analysis

dat <- data %>%
  dplyr::select(id,group, 
                parent_total_verbalisations_time_1, 
                parent_total_verbalisations_time_2, 
                parent_total_verbalisations_time_3) %>%
  pivot_longer( cols = starts_with("parent_total_verbalisations_time_"), 
                names_prefix = "time", 
                names_to="t", values_to="verbal") %>%
  mutate(time = ifelse(t == "parent_total_verbalisations_time_1",  1,
                ifelse(t == "parent_total_verbalisations_time_2", 2, 3))) 
dat1 <- data %>%
  dplyr::select(id,group, 
                parent_total_lp_time_1,
                parent_total_lp_time_2,
                parent_total_lp_time_3) %>%
  pivot_longer( cols = starts_with("parent_total_lp_time_"), 
                names_prefix = "time", 
                names_to="t", values_to="lp") %>%
  mutate(time = ifelse(t == "parent_total_lp_time_1",  1,
                ifelse(t == "parent_total_lp_time_2", 2, 3))) 


p <- left_join(x=dat,y=dat1, by=c("id","time","group")) 

p$time<- as.factor(p$time)
p$group<- as.factor(p$group)


result0<-glmer(cbind(lp, verbal-lp)  ~ (1 | id), data=p, family=binomial)
result1<-glmer(cbind(lp, verbal-lp)  ~ group + (1 | id), data=p, family=binomial)
result2<-glmer(cbind(lp, verbal-lp)  ~ group + time + (1 | id), data=p, family=binomial)
result3<-glmer(cbind(lp, verbal-lp)  ~ group + time +group*time + (1 | id), data=p, family=binomial)

```


This is the ANOVA table for the model. p-values are likelihood ratio tests for adding the new term to the model just before. A p-value <0.05 indicates that adding that term results in better model fit.

```{r, echo=FALSE}
anovatab<-anova(result0,result1,result2,result3)

anovatab <-anovatab %>%
           mutate(Effect = ifelse(npar == 2, "Intercept",
                           ifelse(npar == 4, "Group",
                           ifelse(npar == 6, "Time",
                           ifelse(npar == 9, "Group*Time",""))))) %>%
           relocate(Effect) 
anovatab %>%
           kbl() %>%
           kable_styling()
```



```{r}
summary(result3)
```



#### Estimated Means & Change

Estimated means for each Treatment and Time are presented. In addition, within-group
change between time points is presented with 95% CIs.



```{r, echo=FALSE, warning=FALSE}

result.change  <- emmeans(result3, "time", by = "group", type="response")
result.change.d <-pairs(result.change, reverse=TRUE)
result.change.d.adj <-pairs(result.change, reverse=TRUE,adjust="bonferroni")

# save summary of differences as dataset
estmeans<-summary(result.change)
s<-summary(result.change.d)
ci<-confint(result.change.d)
sadj<-summary(result.change.d.adj)
sadj <- sadj %>%
        rename(p.value.adj=p.value)

s<- left_join(s,sadj)
s<- left_join(s,ci)
```


 Adjusted p-value from Bonferroni correction.

```{r, echo=FALSE}
result.mean  <- emmeans(result3, "group", by = "time", type = "response")
result.mean.d <-pairs(result.mean, type = "response" )
result.mean.d.adj <-pairs(result.mean, adjust="bonferroni")

# save summary of differences as dataset
s_groupdiffs<-summary(result.mean.d)
ci<-confint(result.mean.d)
s_groupdiffsadj<-summary(result.mean.d.adj)
s_groupdiffsadj <- s_groupdiffsadj %>%
        rename(p.value.adj=p.value)

s_groupdiffs<- left_join(s_groupdiffs,s_groupdiffsadj)
s_groupdiffs<- left_join(s_groupdiffs,ci)


```

##### Group mean  at each timepoint


```{r, echo=F}
estmeans %>%
  dplyr::select(c(-df)) %>%
  kbl(digits=5,caption = "Estimated Proportion Number of LP", 
  col.names = c("Time", "Group","Mean", "SE", "LowerCL","UpperCL") )  %>%
  kable_styling() 
```


##### Group mean difference at each timepoint

```{r, echo=F}
s_groupdiffs %>%
dplyr::select(c(-null,-df)) %>%
  kbl(digits=5,caption = "Between-Group Differences", 
  col.names = c("Comparison", "Time","Ratio", "SE", "z","p","adj p",
                  "Lower CL","Upper CL") )  %>%
  kable_styling() %>%
  add_footnote( c("Effect size here is the odds ratio. A value greater than 1 indicates a relative change in odds of LP from the reference group." ), notation = "symbol")
```

Here you can see that Group1 has a lower odds of LP at time 1 than Group2 (OR=.737), but at Time 2, it is much higher (OR=12.808). 


##### Change within each group

```{r, echo=F}

s %>%
 dplyr::select(c(-null,-df)) %>%
  kbl(digits=3,caption = "Within-Group Change", 
  col.names = c("Change", "Group","OR", "SE", "z","p","adj p",
                  "Lower CL","Upper CL") )  %>%
  kable_styling() %>%
  add_footnote( c("Effect size here is the OR of compliance. A value greater than 1 indicates a greater odds of compliance from the reference timepoint." ), notation = "symbol")

```

Here, you can see that the OR for change in group1 from T1 to T2 is 26.55. This can be interpreted as: the odds of LP at time 2 was 26.55 times that of LP at time1. That's huge!  The value for group2 is 1.53, showing an increase in 1.53 times the odds of LP at time 1, but it is so much smaller than the relative increase for Group1.  


```{r, echo=FALSE, warning=FALSE}

result.emm <- emmeans(result3, ~group * time, type = "response")
result.means <- as.data.frame(result.emm)
result.means <-  
  result.means %>%
  mutate(time1= ifelse(group==1 & time==1,  .9,   ## jitter on x axis
                ifelse(group==1 & time==2, 1.9,       
                ifelse(group==1 & time==3, 2.9,       
                ifelse(group==2 & time==1, 1.0,
                ifelse(group==2 & time==2, 2.0,       
                ifelse(group==2 & time==3, 3.0,
                ifelse(group==3 & time==1, 1.1,
                ifelse(group==3 & time==2, 2.1,       
                3.1)))))))))


lp1 <- ggplot(data=result.means, aes(x=time1, y=prob, group=group, shape=group, colour=group)) + geom_line(aes(linetype=group)) + geom_point(size=2) +
    geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL, color=group)) +
  #must define these the same so that only 1 legend is shown
  scale_colour_discrete(name  ="Group",
                            breaks=c(1,2,3),
                            labels=c("PCIT-T", "CoS-P","Waitlist")) +
  scale_shape_discrete(name  ="Group",
                            breaks=c(1,2,3),
                            labels=c("PCIT-T", "CoS-P","Waitlist")) +   
  scale_linetype_manual(name="Group",
                      values =c("solid","dashed","dotted"),
                      breaks=c(1,2,3),
                      labels=c("PCIT-T", "CoS-P","Waitlist")) +
scale_x_continuous("Measurement",breaks = c(1, 2, 3)) +
  scale_y_continuous("Estimated Means (95% CI)") +
theme_classic() 

lp1
  
```

















####  Hypothesis 1

We hypothesize that from baseline (‘T1’) to post-treatment/waitlist (‘T2’) there will be a gradient effect, whereby parent-child dyads who receive PCIT-T
will show the greatest gains and parent-child dyads who
receive no treatment will show the least gains. For the outcome, then, the 
null hypothesis is:

$H_0:  \text {mean change}(\text{PCIT-T}) = \text {mean change}(\text{CoS-P}) = \text {mean change}(\text{Waitlist})$
$H_A: \text {mean change}(\text{PCIT-T}) > \text {mean change}(\text{CoS-P}) > \text {mean change}(\text{Waitlist})$


This is a joint hypothesis test.

First, confirming that the ordering of the means is as hypothesized:

```{r, echo=FALSE}
confint(result.change.d)
# df= Inf is ok
```

Now, to get the p-value associated with the joint hypothesis.

```{r}
## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp1   <- matrix(c(0, 0,0, 0,0, -1,0,0 ),1)
# Hyp2 is 1 df test, 
hyp2   <- matrix(c(0, 0,0, 0,1, 0,1,0 ),1)


# hypF is combined 2df test. Yes!
h<-rbind(hyp1,hyp2)
hypF <- as.matrix(h)
GlobalF <- glht(result3, linfct = hypF, test = Ftest())
GlobalF$linfct
summary(GlobalF, test = Ftest())

#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html

```

p<0.05 so this hypothesis is supported.


#### Hypothesis 2


##### Change from T2-T3

There will be a gradient of size of change from T2 to T3 such that the change in the PCIT-T group > the change in the CoS-P group.  This isn't really a gradient; it's just 1 hypothesis.

First, the change in PCITT and COSP from time 2 to time 3:

```{r, echo=FALSE}
result.change

      
confint(result.change.d)
or<-data.frame(result.change.d)
```




```{r}
## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, -1,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2)
summary(hyp2F)
s<-summary(hyp2F, test = Ftest())
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html



## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```

Change in G1 from T2 to T3: `r round(or[3,3],3)`

Change in G2 from T2 to T3: `r round(or[6,3],3)`

p=`r round(s$test$pvalue,4)`. We have `r evidence` that the change from T1 to T3 is different for G1 and G2 


##### Change from T1-T3



```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, 0,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2, test = Ftest())
summary(hyp2F, test = Ftest())
s<-summary(hyp2F, test = Ftest())
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html


## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```


Change in G1 from T1 to T3: `r round(or[2,3],3)`

Change in G2 from T1 to T3: `r round(or[5,3],3)`


p=`r round(s$test$pvalue,4)`. We have `r evidence` that the change from T1 to T3 is different for G1 and G2 



#### Residuals


Before interpretation, we should examine model assumptions.

For this model (A generalized linear mixed model with poisson distribution and log link, with random effect for individual), we have to use a different residual than for a linear mixed model. These are simulated residuals, using the DHARMa package in R (citation below).

Under H0 (perfect model, no violations of assumptions), we would expect a straight line in the QQplot. Also, scatterplot should show similar spread across the x-axis.

There are no violations of assumptions.

```{r}

library(DHARMa)

sim= simulateResiduals(result3)
resid = residuals(sim, quantileFunction = qnorm, outlierValues = c(-7,7))
plot(resid~predict(result3, re.form = NA,type="response"))
qqnorm(resid)

citation("DHARMa")
```



```{r}
rm(list = ls())
```