---
title: "Hypothesis Tests"
author: "Nancy Briggs"
date: "`r Sys.Date()`"
output: html_document
---


#### Hypothesis 1

We hypothesize that from baseline (‘T1’) to post-treatment/waitlist (‘T2’) there will be a gradient effect, whereby parent-child dyads who receive PCIT-T
will show the greatest gains and parent-child dyads who
receive no treatment will show the least gains. For the outcome, then, the 
null hypothesis is:


$H_0:  \text {mean change}(\text{PCIT-T}) = \text {mean change}(\text{CoS-P}) = \text {mean change}(\text{Waitlist})$
$H_A: \text {mean change}(\text{PCIT-T}) > \text {mean change}(\text{CoS-P}) > \text {mean change}(\text{Waitlist})$

This is a joint hypothesis test.

First, confirming that the ordering of the means is as hypothesized:

```{r, echo=FALSE}
result.change.d
```

So, the mean change for PCITT > COSP > Waitlist.

Now, to get the p-value associated with the joint hypothesis.


```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp1   <- matrix(c(0, 0,0, 0,0, -1,0,0 ),1)
# Hyp2 is 1 df test, 
hyp2   <- matrix(c(0, 0,0, 0,1, 0,1,0 ),1)

# hypF is combined 2df test. Yes!
h<-rbind(hyp1,hyp2)
hypF <- as.matrix(h)
GlobalF <- glht(result3, linfct = hypF, test = Ftest())
GlobalF$linfct
summary(GlobalF, test = Ftest())
s<-summary(GlobalF, test = Ftest())

#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html

## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```

p=`r round(s$test$pvalue,4)`. We have `r evidence` that this the change from T1 to T2 is greater for G1 then G2 and the change in G2 is greater than in G3.





#### Hypothesis 2

There will be a gradient of size of change from T2 to T3 such that the change in the PCIT-T group > the change in the CoS-P group.  This isn't really a gradient; it's just 1 hypothesis.

First, the change in PCITT and COSP from time 2 to time 3:

```{r, echo=FALSE}

result.change.d1<-as.data.frame(result.change.d)
result.change.d1<- result.change.d1 %>%
  mutate(t2t3change=-1*estimate) %>%
  mutate(quality=ifelse(t2t3change<0, "decreased t2 to t3", 
                 ifelse(t2t3change>0, "increase t2 to t3", "did not change" )))

result.change.d1 %>%
  filter(contrast=="2 - 3") %>%
  dplyr::select(contrast, group, t2t3change, quality, p.value ) %>%
  kbl(digits=3) %>%
  kable_styling()

```



```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, -1,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2)
summary(hyp2F)
F<-summary(hyp2F)
ci<- confint(F)
ci
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html

## Extract pvaue for evidence / no evidence
evidence = ifelse(F$test$pvalue<0.05, "evidence", "no evidence")

```

p=`r round(F$test$pvalue,3)`, so we have `r evidence` to suggest that the change from t2 to t3 in PCITT is different than the change in COSP.




```{r, echo=FALSE, warning=FALSE}
# ordering change.
## Extract pvalue for evidence / no evidence
orderincrease <- ifelse(result.change.d1$t2t3change[3]>0 &


result.change.d1$t2t3change[3]>result.change.d1$t2t3change[6],
                 "PCITT > COS-P, increase was greater in PCITT", 
                 "PCITT < COS-P, increase was not greater in PCITT" )
orderdecrease <- ifelse(result.change.d1$t2t3change[3]<0 &
                        result.change.d1$t2t3change[3]<result.change.d1$t2t3change[6],
                 "PCITT > COS-P, decrease was greater in PCITT", 
                 "PCITT < COS-P, decrease was not greater in PCITT" )

```

Assuming an increase in scores is better, `r orderincrease`

Assuming a decrease in scores is better,`r orderdecrease`


