---
title: "PCIT-CoS-Waitlist RCT Paper 1: DPICS Child Compliance"
author: "Nancy Briggs"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---


### DPICS Child Compliance  {.tabset .tabset-fade .tabset-pills}

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)
options(warn=1)

library(lme4)
library(tidyverse)
library(tidyselect)
library(SASmarkdown)
library(nlme)
library(emmeans)
library(multcomp)
library(kableExtra)
library(ggplot2)
library(lmerTest)
library(DHARMa)
library(effectsize)
library(boot)

```


```{r, echo=FALSE}
# Data 
# Import data
# recoding for CBCL and totals done in SPSS before bringing data in to R

data <- read.csv("../data_prep.csv", fileEncoding="UTF-8-BOM")
names(data) <- tolower(names(data))

# keep relevant demographic and outcome variables

data <- data %>%
  dplyr::select(id:treatment_time_mins, dobchild, sexchild, numbhousehold, incfamily,
                
                #Child Compliance
                total_co_perc_time_1,total_co_perc_time_2,total_co_perc_time_3
                
                # others, but not essential for paper
                #3. Child Temperament (ECBQ)
                #4. Child social-emotional functioning - #1: Brief Infant Toddler Social Emotional Assessment (BITSEA)
                #5. Child social-emotional functioning - #2: Child Devereux Early Childhood Assessment (DECA)
                
                )


### import total time
time <- read.csv("../data_prep_DPiCStime.csv", fileEncoding="UTF-8-BOM")
names(time) <- tolower(names(time))
time <- time %>%
  dplyr::select(id, parent_total_verbalisations_time_1, parent_total_lp_time_1, 
                  parent_total_nta_time_1, total_eff_commands_time_1, 
                  child_co_total_time_1, child_nc_total_time_1, 
                  total_nc_perc_time_1, 
                  parent_total_verbalisations_time_2, parent_total_lp_time_2, 
                  parent_total_nta_time_2, total_eff_commands_time_2, 
                  child_co_total_time_2, child_nc_total_time_2, 
                  total_nc_perc_time_2, 
                  parent_total_verbalisations_time_3, parent_total_lp_time_3, 
                  parent_total_nta_time_3, total_eff_commands_time_3, 
                  child_co_total_time_3, child_nc_total_time_3, 
                  total_nc_perc_time_3	)

# Join
data <- left_join(data, time, by="id")


# there is one case in g3 waitlist that has infoor for T3.  recode compliance 
# for that person to NA.
data$total_co_perc_time_3[data$group==3] <- NA
data$child_co_total_time_3[data$group==3] <- NA
```




#### Descriptives

_Descriptives for Percentage of Effective Commands_

```{r, echo=FALSE, warning=FALSE}
#Child Compliance
#total_co_perc_time_1,total_co_perc_time_2,total_co_perc_time_3
# One person on Waitlist has a value for FU.  set to NA
data$total_co_perc_time_3[data$group==3] <- NA
         
desctable <- data %>%
  group_by(group) %>%
  mutate(score1 = case_when( 
    total_co_perc_time_1 == -9 ~ NA_real_,   #assign missing values to zeros
    TRUE ~ total_co_perc_time_1)) %>%
  mutate(score2 = case_when( 
    total_co_perc_time_2 == -9 ~ NA_real_,    #assign missing values to zeros
    TRUE ~ total_co_perc_time_2)) %>%
  mutate(score3 = case_when( 
    total_co_perc_time_3 == -9 ~ NA_real_,   #assign missing values to zeros
    TRUE ~ total_co_perc_time_3)) %>%
  summarise(n1 = sum(!is.na(score1)) , 
            mean1=mean(total_co_perc_time_1, na.rm = T ),
            sd1=sd(total_co_perc_time_1, na.rm = T ),
            median1=median(total_co_perc_time_1, na.rm = T ),
            min1=min(total_co_perc_time_1, na.rm = T ),
            max1=max(total_co_perc_time_1, na.rm = T ),
            
            n2 = sum(!is.na(score2)) , 
            mean2=mean(total_co_perc_time_2, na.rm = T ),
            sd2=sd(total_co_perc_time_2, na.rm = T ),
            median2=median(total_co_perc_time_2, na.rm = T ),
            min2=min(total_co_perc_time_2, na.rm = T ),
            max2=max(total_co_perc_time_2, na.rm = T ),
            
            n3 = sum(!is.na(score3)) , 
            mean3=mean(total_co_perc_time_3, na.rm = T ),
            sd3=sd(total_co_perc_time_3, na.rm = T ),
            median3=median(total_co_perc_time_3, na.rm = T ),
            min3=min(total_co_perc_time_3, na.rm = T ),
            max3=max(total_co_perc_time_3, na.rm = T )  ) %>%
  pivot_longer(cols = -group,  
                names_to = c('.value', 'time'), 
                    names_pattern = '(.*)(\\d+)') 
  


write.csv(desctable,"desctable.csv", row.names = TRUE)

```


```{r, echo=FALSE}
desctable %>%
           kbl() %>%
           kable_styling() %>%
           footnote("One person in the Waitlist group had a value at Follow Up. It was set to missing.")
```



```{r, echo=FALSE}
#Child Compliance
#total_co_perc_time_1,total_co_perc_time_2,total_co_perc_time_3

# Histograms
# plot
p <- data %>%
  dplyr::select(id,group, total_co_perc_time_1, total_co_perc_time_2, total_co_perc_time_3) %>%
  pivot_longer( cols = starts_with("total_co_perc_time_"), names_prefix = "time", 
                names_to="t", values_to="outcome") %>%
  mutate(time = ifelse(t == "total_co_perc_time_1",  1,
                ifelse(t == "total_co_perc_time_2", 2, 3))) 
  

  
hist<-ggplot(p, aes(x=outcome) )+
  geom_histogram(alpha=0.6, binwidth = 5) +
  xlab("Child Compliance Percentage") +
  ylab("Frequency") +
  facet_wrap(group~time)

group.labs <- c("PCIT-T", "COS-P", "Waitlist")
names(group.labs) <- c("1", "2", "3")

# New facet label names for group variable
time.labs <- c("Pre", "Post","FollowUp")
names(time.labs) <- c("1", "2","3")
# Create the plot
hist + theme_minimal() +
  facet_grid(
  group ~ time, 
  labeller = labeller(group = group.labs, time = time.labs) )

```


#### Model & Results

A generalized linear mixed model (GLMM) with random effect for individual and fixed effects of group, time and the interaction was run on the number of Compliance responses in the total number of verbalizations. A binomial distribution was specified with a logit link. Estimated marginal means were obtained, and contrasts specified to test Hypothesis 1 and Hypothesis 2. Assumptions on the residuals were examined using simulated residuals (see below).

TOTAL_eff_commands_TIME_1 is the denominator, and CHILD_CO_total_TIME_1 is the numerator


```{r, echo=FALSE}
#  analysis
#TOTAL_eff_commands_TIME_1
#CHILD_CO_total_TIME_1

dat <- data %>%
  dplyr::select(id,group, 
                total_eff_commands_time_1, 
                total_eff_commands_time_2, 
                total_eff_commands_time_3) %>%
  pivot_longer( cols = starts_with("total_eff_commands_time_"), 
                names_prefix = "time", 
                names_to="t", values_to="effcommands") %>%
  mutate(time = ifelse(t == "total_eff_commands_time_1",  1,
                ifelse(t == "total_eff_commands_time_2", 2, 3))) 
dat1 <- data %>%
  dplyr::select(id,group, 
                child_co_total_time_1,
                child_co_total_time_2,
                child_co_total_time_3) %>%
  pivot_longer( cols = starts_with("child_co_total_time_"), 
                names_prefix = "time", 
                names_to="t", values_to="co") %>%
  mutate(time = ifelse(t == "child_co_total_time_1",  1,
                ifelse(t == "child_co_total_time_2", 2, 3))) 


p <- left_join(x=dat,y=dat1, by=c("id","time","group")) 

p$time<- as.factor(p$time)
p$group<- as.factor(p$group)


result0<-glmer(cbind(co, effcommands-co) ~ (1 | id), data=p, family=binomial)
result1<-glmer(cbind(co, effcommands-co) ~ group + (1 | id), data=p, family=binomial)
result2<-glmer(cbind(co, effcommands-co) ~ group + time + (1 | id), data=p, family=binomial)
result3<-glmer(cbind(co, effcommands-co) ~ group + time +group*time + (1 | id), data=p, family=binomial)

```


This is the ANOVA table for the model. p-values are likelihood ratio tests for adding the new term to the model just before. A p-value <0.05 indicates that adding that term results in better model fit.

```{r, echo=FALSE}
anovatab<-anova(result0,result1,result2,result3)

anovatab <-anovatab %>%
           mutate(Effect = ifelse(npar == 2, "Intercept",
                           ifelse(npar == 4, "Group",
                           ifelse(npar == 6, "Time",
                           ifelse(npar == 9, "Group*Time",""))))) %>%
           relocate(Effect) 
anovatab %>%
           kbl() %>%
           kable_styling()
```



```{r}
summary(result3)
```



#### Estimated Means & Change


```{r, echo=FALSE, warning=FALSE}

result.change  <- emmeans(result3, "time", by = "group", type="response")
result.change.d <-pairs(result.change, reverse=TRUE)
result.change.d.adj <-pairs(result.change, reverse=TRUE,adjust="bonferroni")

# save summary of differences as dataset
estmeans<-summary(result.change)
s<-summary(result.change.d)
ci<-confint(result.change.d)
sadj<-summary(result.change.d.adj)
sadj <- sadj %>%
        rename(p.value.adj=p.value)

s<- left_join(s,sadj)
s<- left_join(s,ci)
```


 Adjusted p-value from Bonferroni correction.

```{r, echo=FALSE}
result.mean  <- emmeans(result3, "group", by = "time", type = "response")
result.mean.d <-pairs(result.mean, type = "response" )
result.mean.d.adj <-pairs(result.mean, adjust="bonferroni")

# save summary of differences as dataset
s_groupdiffs<-summary(result.mean.d)
ci<-confint(result.mean.d)
s_groupdiffsadj<-summary(result.mean.d.adj)
s_groupdiffsadj <- s_groupdiffsadj %>%
        rename(p.value.adj=p.value)

s_groupdiffs<- left_join(s_groupdiffs,s_groupdiffsadj)
s_groupdiffs<- left_join(s_groupdiffs,ci)


```

##### Group mean  at each timepoint


```{r, echo=F}
estmeans %>%
  dplyr::select(c(-df)) %>%
  kbl(digits=3,caption = "Estimated Mean Number of LP", 
  col.names = c("Time", "Group","Mean", "SE", "LowerCL","UpperCL") )  %>%
  kable_styling() 
```


##### Group mean difference at each timepoint

```{r, echo=F}
s_groupdiffs %>%
dplyr::select(c(-null,-df)) %>%
  kbl(digits=3,caption = "Between-Group Differences", 
  col.names = c("Comparison", "Time","Ratio", "SE", "z","p","adj p",
                  "Lower CL","Upper CL") )  %>%
  kable_styling() %>%
  add_footnote( c("Effect size here is the ratio of means A value greater than 1 indicates a proportion change from the reference group." ), notation = "symbol")
```

Here you can see that group1 shows higher odds of compliance than group2 at time 2 (OR=1.157), but the CI is wide (0.695, 1.926) and the p>0.05.


##### Change within each group

```{r, echo=F}

s %>%
 dplyr::select(c(-null,-df)) %>%
  kbl(digits=3,caption = "Within-Group Change", 
  col.names = c("Change", "Group","Ratio", "SE", "z","p","adj p",
                  "Lower CL","Upper CL") )  %>%
  kable_styling() %>%
  add_footnote( c("Effect size here is the ratio of change. A value greater than 1 indicates a proportion change from the reference timepoint." ), notation = "symbol")

```

Here, you can see that the OR for change in group1 from T1 to T2 is 1.357 (-/986, 1.968) and the OR for change in group2 is 1.210 (0.861, 1.701). Both of these do show an increase in odds of compliance over time, but the change is not significant.

```{r, echo=FALSE, warning=FALSE}

result.emm <- emmeans(result3, ~group * time, type = "response")
result.means <- as.data.frame(result.emm)
result.means <-  
  result.means %>%
  mutate(time1= ifelse(group==1 & time==1,  .9,   ## jitter on x axis
                ifelse(group==1 & time==2, 1.9,       
                ifelse(group==1 & time==3, 2.9,       
                ifelse(group==2 & time==1, 1.0,
                ifelse(group==2 & time==2, 2.0,       
                ifelse(group==2 & time==3, 3.0,
                ifelse(group==3 & time==1, 1.1,
                ifelse(group==3 & time==2, 2.1,       
                3.1)))))))))


lp1 <- ggplot(data=result.means, aes(x=time1, y=prob, group=group, shape=group, colour=group)) + geom_line(aes(linetype=group)) + geom_point(size=2) +
    geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL, color=group)) +
  #must define these the same so that only 1 legend is shown
  scale_colour_discrete(name  ="Group",
                            breaks=c(1,2,3),
                            labels=c("PCIT-T", "CoS-P","Waitlist")) +
  scale_shape_discrete(name  ="Group",
                            breaks=c(1,2,3),
                            labels=c("PCIT-T", "CoS-P","Waitlist")) +   
  scale_linetype_manual(name="Group",
                      values =c("solid","dashed","dotted"),
                      breaks=c(1,2,3),
                      labels=c("PCIT-T", "CoS-P","Waitlist")) +
scale_x_continuous("Measurement",breaks = c(1, 2, 3)) +
  scale_y_continuous("Estimated Means (95% CI)") +
theme_classic() 

lp1
  
```















####  Hypothesis 1

We hypothesize that from baseline (‘T1’) to post-treatment/waitlist (‘T2’) there will be a gradient effect, whereby parent-child dyads who receive PCIT-T
will show the greatest gains and parent-child dyads who
receive no treatment will show the least gains. For the outcome, then, the 
null hypothesis is:

$H_0:  \text {mean change}(\text{PCIT-T}) = \text {mean change}(\text{CoS-P}) = \text {mean change}(\text{Waitlist})$
$H_A: \text {mean change}(\text{PCIT-T}) > \text {mean change}(\text{CoS-P}) > \text {mean change}(\text{Waitlist})$


This is a joint hypothesis test.

First, confirming that the orderig of the means is as hypothesized:

```{r, echo=FALSE}
confint(result.change.d)
# df= Inf is ok
```


Here, the odds of compliance 1.36 times the odds at time 2 than time 1 for PCIT-T.  Or, there is a 36% increase in the odds of compliance at Time 2.

Now, to get the p-value associated with the joint hypothesis.

```{r}
## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp1   <- matrix(c(0, 0,0, 0,0, -1,0,0 ),1)
# Hyp2 is 1 df test, 
hyp2   <- matrix(c(0, 0,0, 0,1, 0,1,0 ),1)


# hypF is combined 2df test. Yes!
h<-rbind(hyp1,hyp2)
hypF <- as.matrix(h)
GlobalF <- glht(result3, linfct = hypF, test = Ftest())
GlobalF$linfct
summary(GlobalF, test = Ftest())

#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html

```

p>0.05 so this hypothesis is not supported. 



#### Hypothesis 2

##### Change from T2-T3

There will be a gradient of size of change from T2 to T3 such that the change in the PCIT-T group > the change in the CoS-P group.  This isn't really a gradient; it's just 1 hypothesis.

First, the change in PCITT and COSP from time 2 to time 3:

```{r, echo=FALSE}
result.change

      
confint(result.change.d)
or<-data.frame(result.change.d)
```




```{r}
## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, -1,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2)
summary(hyp2F)
s<-summary(hyp2F, test = Ftest())
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html



## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```

Change in G1 from T2 to T3: `r round(or[3,3],3)`

Change in G2 from T2 to T3: `r round(or[6,3],3)`

p=`r round(s$test$pvalue,4)`. We have `r evidence` that the change from T1 to T3 is different for G1 and G2 


##### Change from T1-T3



```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, 0,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2, test = Ftest())
summary(hyp2F, test = Ftest())
s<-summary(hyp2F, test = Ftest())
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html


## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```


Change in G1 from T1 to T3: `r round(or[2,3],3)`

Change in G2 from T1 to T3: `r round(or[5,3],3)`


p=`r round(s$test$pvalue,4)`. We have `r evidence` that the change from T1 to T3 is different for G1 and G2 


#### Residuals


Before interpretation, we should examine model assumptions.

For this model (A generalized linear mixed model with binomial distribution and logit link, with random effect for individual), we have to use a different residual than for a linear mixed model. These are simulated residuals, using the DHARMa package in R (citation below).

Under H0 (perfect model, no violations of assumptions), we would expect a straight line in the QQplot. Also, the boxes in the boxplot should range homogenously from 0.25-0.75. To see whether there are deviations from this expectation, the plot calculates a test for uniformity per box, and a test for homgeneity of variances between boxes. 

There are no violations of assumptions.

```{r}

library(DHARMa)

sim= simulateResiduals(result3)
resid = residuals(sim, quantileFunction = qnorm, outlierValues = c(-7,7))
plot(resid~predict(result3, re.form = NA,type="response"))
qqnorm(resid)

citation("DHARMa")
```


There are a 3 observations that are not fit well  by the model.  These observations show very high or very low proportion of compliance to effective commands.  


#### Sensitivity

Here is a quick sensitivity check excluding those 3 cases with very different proportions

```{r, echo=F}
filter<-c(91,17,54)
p %>%
  filter(id %in% filter)

p1 <- p %>%
  mutate(co=ifelse(id==91 & time==1,NA,co)) %>%
  mutate(co=ifelse(id==17 & time==3,NA,co)) %>%
  mutate(co=ifelse(id==54 & time==2,NA,co))


```

You can see that for id 91, the time 1 value is relatively high (28/34). For id 17 the time 3 value is low (2/20) and for id 54, the time 2 value is low (4/41). These observations will be removed and the model rerun.

```{r, echo=F}
result0<-glmer(cbind(co, effcommands-co) ~ (1 | id), data=p1, family=binomial)
result1<-glmer(cbind(co, effcommands-co) ~ group + (1 | id), data=p1, family=binomial)
result2<-glmer(cbind(co, effcommands-co) ~ group + time + (1 | id), data=p1, family=binomial)
result3<-glmer(cbind(co, effcommands-co) ~ group + time +group*time + (1 | id), data=p1, family=binomial)
```

```{r, echo=F}
anovatab<-anova(result0,result1,result2,result3)

anovatab <-anovatab %>%
           mutate(Effect = ifelse(npar == 2, "Intercept",
                           ifelse(npar == 4, "Group",
                           ifelse(npar == 6, "Time",
                           ifelse(npar == 9, "Group*Time",""))))) %>%
           relocate(Effect) 
anovatab %>%
           kbl() %>%
           kable_styling()
```
group*time effect is still not significant.


```{r, echo=F}
summary(result3)


```



#### Estimated Means & Change


```{r, echo=FALSE, warning=FALSE}

result.change  <- emmeans(result3, "time", by = "group", type="response")
result.change.d <-pairs(result.change, reverse=TRUE)
result.change.d.adj <-pairs(result.change, reverse=TRUE,adjust="bonferroni")

# save summary of differences as dataset
estmeans<-summary(result.change)
s<-summary(result.change.d)
ci<-confint(result.change.d)
sadj<-summary(result.change.d.adj)
sadj <- sadj %>%
        rename(p.value.adj=p.value)

s<- left_join(s,sadj)
s<- left_join(s,ci)
```


 Adjusted p-value from Bonferroni correction.

```{r, echo=FALSE}
result.mean  <- emmeans(result3, "group", by = "time", type = "response")
result.mean.d <-pairs(result.mean, type = "response" )
result.mean.d.adj <-pairs(result.mean, adjust="bonferroni")

# save summary of differences as dataset
s_groupdiffs<-summary(result.mean.d)
ci<-confint(result.mean.d)
s_groupdiffsadj<-summary(result.mean.d.adj)
s_groupdiffsadj <- s_groupdiffsadj %>%
        rename(p.value.adj=p.value)

s_groupdiffs<- left_join(s_groupdiffs,s_groupdiffsadj)
s_groupdiffs<- left_join(s_groupdiffs,ci)


```

##### Group mean  at each timepoint


```{r, echo=F}
estmeans %>%
  dplyr::select(c(-df)) %>%
  kbl(digits=3,caption = "Estimated Mean Number of LP", 
  col.names = c("Time", "Group","Mean", "SE", "LowerCL","UpperCL") )  %>%
  kable_styling() 
```


##### Group mean difference at each timepoint

```{r, echo=F}
s_groupdiffs %>%
dplyr::select(c(-null,-df)) %>%
  kbl(digits=3,caption = "Between-Group Differences", 
  col.names = c("Comparison", "Time","Ratio", "SE", "z","p","adj p",
                  "Lower CL","Upper CL") )  %>%
  kable_styling() %>%
  add_footnote( c("Effect size here is the ratio of means A value greater than 1 indicates a proportion change from the reference group." ), notation = "symbol")
```
Nothing changed from the previous model.


##### Change within each group

```{r, echo=F}

s %>%
 dplyr::select(c(-null,-df)) %>%
  kbl(digits=3,caption = "Within-Group Change", 
  col.names = c("Change", "Group","Ratio", "SE", "z","p","adj p",
                  "Lower CL","Upper CL") )  %>%
  kable_styling() %>%
  add_footnote( c("Effect size here is the ratio of change. A value greater than 1 indicates a proportion change from the reference timepoint." ), notation = "symbol")

```

Nothing changed from the previous model.

```{r, echo=FALSE, warning=FALSE}

result.emm <- emmeans(result3, ~group * time, type = "response")
result.means <- as.data.frame(result.emm)
result.means <-  
  result.means %>%
  mutate(time1= ifelse(group==1 & time==1,  .9,   ## jitter on x axis
                ifelse(group==1 & time==2, 1.9,       
                ifelse(group==1 & time==3, 2.9,       
                ifelse(group==2 & time==1, 1.0,
                ifelse(group==2 & time==2, 2.0,       
                ifelse(group==2 & time==3, 3.0,
                ifelse(group==3 & time==1, 1.1,
                ifelse(group==3 & time==2, 2.1,       
                3.1)))))))))


lp1 <- ggplot(data=result.means, aes(x=time1, y=prob, group=group, shape=group, colour=group)) + geom_line(aes(linetype=group)) + geom_point(size=2) +
    geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL, color=group)) +
  #must define these the same so that only 1 legend is shown
  scale_colour_discrete(name  ="Group",
                            breaks=c(1,2,3),
                            labels=c("PCIT-T", "CoS-P","Waitlist")) +
  scale_shape_discrete(name  ="Group",
                            breaks=c(1,2,3),
                            labels=c("PCIT-T", "CoS-P","Waitlist")) +   
  scale_linetype_manual(name="Group",
                      values =c("solid","dashed","dotted"),
                      breaks=c(1,2,3),
                      labels=c("PCIT-T", "CoS-P","Waitlist")) +
scale_x_continuous("Measurement",breaks = c(1, 2, 3)) +
  scale_y_continuous("Estimated Means (95% CI)") +
theme_classic() 

lp1
  
```



Takeaway: removing those relatively extreme values doesn't change conclusions.




```{r}
rm(list = ls())
```