---
title: "PCIT-CoS-Waitlist RCT Paper 1: Observed maternal sensitivity (NICHD)"
author: "Nancy Briggs"
date: "06 June 2022"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---


### Observed maternal sensitivity (NICHD) {.tabset .tabset-fade .tabset-pills}

Output run on `r Sys.Date()`

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)

library(tidyverse)
library(tidyselect)
library(SASmarkdown)
library(nlme)
library(emmeans)
library(multcomp)
library(kableExtra)
library(ggplot2)
library(lmerTest)
library(effectsize)
library(boot)

options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)

```


```{r, echo=FALSE}
# Data 
# Import data
# recoding for CBCL and totals done in SPSS before bringing data in to R

data <- read.csv("../data_prep.csv", fileEncoding="UTF-8-BOM")
names(data) <- tolower(names(data))

# keep relevant demographic and outcome variables

data <- data %>%
  dplyr::select(id:treatment_time_mins, dobchild, sexchild, numbhousehold, incfamily,
 
                #NICHD
                #Observed maternal negativity
                nichd_negative_total, post_nichd_negative_total, fu_nichd_negative_total,
                #Observed maternal sensitivity 
                nichd_total_sensitivity, post_nichd_total_sensitivity, fu_nichd_total_sensitivity

                )


```




#### Descriptives

```{r, echo=FALSE}
# nichd_total_sensitivity, post_nichd_total_sensitivity, fu_nichd_total_sensitivity

# One person on Waitlist has a value for FU.  set to NA
data$fu_nichd_total_sensitivity[data$group==3] <- NA

desctable <- data %>%
  group_by(group) %>%
  summarise(n1 = sum(!is.na(nichd_total_sensitivity)) , 
            mean1=mean(nichd_total_sensitivity, na.rm = T ),
            sd1=sd(nichd_total_sensitivity, na.rm = T ),
            median1=median(nichd_total_sensitivity, na.rm = T ),
            min1=min(nichd_total_sensitivity, na.rm = T ),
            max1=max(nichd_total_sensitivity, na.rm = T ),
            
            n2 = sum(!is.na(post_nichd_total_sensitivity)) , 
            mean2=mean(post_nichd_total_sensitivity, na.rm = T ),
            sd2=sd(post_nichd_total_sensitivity, na.rm = T ),
            median2=median(post_nichd_total_sensitivity, na.rm = T ),
            min2=min(post_nichd_total_sensitivity, na.rm = T ),
            max2=max(post_nichd_total_sensitivity, na.rm = T ),
            
            n3 = sum(!is.na(fu_nichd_total_sensitivity)) , 
            mean3=mean(fu_nichd_total_sensitivity, na.rm = T ),
            sd3=sd(fu_nichd_total_sensitivity, na.rm = T ),
            median3=median(fu_nichd_total_sensitivity, na.rm = T ),
            min3=min(fu_nichd_total_sensitivity, na.rm = T ),
            max3=max(fu_nichd_total_sensitivity, na.rm = T )  ) %>%
  pivot_longer(cols = -group,  
                names_to = c('.value', 'time'), 
                    names_pattern = '(.*)(\\d+)')

write.csv(desctable,"desctable.csv", row.names = TRUE)



```

```{r, echo=FALSE}
desctable %>%
           kbl(digits=3, caption="Raw Data") %>%
           kable_minimal() %>%
           footnote("One person the Waitlist has a value at Follow Up. It has been set to missing.")


```



```{r, echo=FALSE}
##   nichd_total_sensitivity, post_nichd_total_sensitivity, fu_nichd_total_sensitivity
# Histograms
# plot
p <- data %>%
  dplyr::select(id,group,  nichd_total_sensitivity, post_nichd_total_sensitivity, fu_nichd_total_sensitivity) %>%
  pivot_longer( cols = c( "nichd_total_sensitivity", "post_nichd_total_sensitivity", "fu_nichd_total_sensitivity"),  names_prefix = "bitseacomp_", 
                names_to="t", values_to="outcome") %>%
  mutate(time = ifelse(t == "nichd_total_sensitivity",  1,
                ifelse(t == "post_nichd_total_sensitivity", 2, 3))) 
  

  
hist<-ggplot(p, aes(x=outcome) )+
  geom_histogram(alpha=0.6, binwidth = 5) +
  xlab("Observed maternal sensitivity (NICHD)") +
  ylab("Frequency") +
  facet_wrap(group~time)

group.labs <- c("PCIT-T", "COS-P", "Waitlist")
names(group.labs) <- c("1", "2", "3")

# New facet label names for group variable
time.labs <- c("Pre", "Post","FollowUp")
names(time.labs) <- c("1", "2","3")
# Create the plot
hist + theme_minimal() +
  facet_grid(
  group ~ time, 
  labeller = labeller(group = group.labs, time = time.labs) )

```

#### Model & Results

A linear mixed model with random effect for individual and fixed effects of group, time and the interaction was run *on the log-transformed data*. Estimated marginal means were obtained, and contrasts specified to test Hypothesis 1 and Hypothesis 2.

```{r, echo=FALSE, warning=FALSE}
#  analysis
dat <- data %>%
dplyr::select(id,group,  nichd_total_sensitivity, post_nichd_total_sensitivity, fu_nichd_total_sensitivity) %>%
  pivot_longer( cols = c( "nichd_total_sensitivity", "post_nichd_total_sensitivity", "fu_nichd_total_sensitivity"),  names_prefix = "bitseacomp_", 
                names_to="t", values_to="outcome") %>%
  mutate(time = ifelse(t == "nichd_total_sensitivity",  1,
                ifelse(t == "post_nichd_total_sensitivity", 2, 3))) 
  

dat$time<- as.factor(dat$time)
dat$group<- as.factor(dat$group)

result0<-lmer(outcome ~ (1 | id), dat)
result1<-lmer(outcome ~ group + (1 | id), dat)
result2<-lmer(outcome ~ group + time + (1 | id), dat)
result3<-lmer(outcome ~ group*time + (1 | id), dat)

```




This is the ANOVA table for the model. p-values are likelihood ratio tests for adding the new term to the model just before. A p-value <0.05 indicates that adding that term results in better model fit. The most important p-value is the interaction p-value. Here, there is  evidence that the pattern of change over time is different between the groups (p=0.003).

```{r, echo=FALSE}
anovatab<-anova(result0,result1,result2,result3)

anovatab <-anovatab %>%
           mutate(Effect = ifelse(npar == 3, "Intercept",
                           ifelse(npar == 5, "Group",
                           ifelse(npar == 7, "Time",
                           ifelse(npar == 10, "Group*Time",""))))) %>%
           relocate(Effect) 
anovatab %>%
           kbl(digits=3) %>%
           kable_styling()
```



```{r}
summary(result3)
```



#### Estimated Means & Change

Estimated means for each Treatment and Time are presented. In addition, within-group
change between time points is presented with 95% CIs.


```{r child='bootstrapped_es_CI.Rmd'}
```


#### Hypothesis 1

We hypothesize that from baseline (‘T1’) to post-treatment/waitlist (‘T2’) there will be a gradient effect, whereby parent-child dyads who receive PCIT-T
will show the greatest gains and parent-child dyads who
receive no treatment will show the least gains. For the outcome, then, the 
null hypothesis is:

$H_0:  \text {mean change}(\text{PCIT-T}) = \text {mean change}(\text{CoS-P}) = \text {mean change}(\text{Waitlist})$
$H_A: \text {mean change}(\text{PCIT-T}) > \text {mean change}(\text{CoS-P}) > \text {mean change}(\text{Waitlist})$

This is a joint hypothesis test.

First, confirming that the ordering of the means is as hypothesized:

```{r, echo=FALSE}
result.change  <- emmeans(result3, "time", by = "group")
result.change.d <-pairs(result.change)
result.change.d
```

scores increased for the PCITT group from Baseline to post (-10.55), and increased for COSP (-3.83).

```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp1   <- matrix(c(0, 0,0, 0,0, -1,0,0 ),1)
# Hyp2 is 1 df test, 
hyp2   <- matrix(c(0, 0,0, 0,1, 0,1,0 ),1)

# hypF is combined 2df test. Yes!
h<-rbind(hyp1,hyp2)
hypF <- as.matrix(h)
GlobalF <- glht(result3, linfct = hypF, test = Ftest())
GlobalF$linfct
summary(GlobalF, test = Ftest())

#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html

```

pvalue>0.08. We do not have evidence that there is a difference in the change over from Baseline to Post amongst the groups.  My guess is the individual test of difference begtween PCITT and COSP is significant.




#### Hypothesis 2

##### Change from T2-T3

There will be a gradient of size of change from T2 to T3 such that the change in the PCIT-T group > the change in the CoS-P group.  This isn't really a gradient; it's just 1 hypothesis.

First, the change in PCITT and COSP from time 2 to time 3:

```{r, echo=FALSE}
result.change

      
confint(result.change.d)
or<-data.frame(result.change.d)
```




```{r}
## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, -1,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2)
summary(hyp2F)
s<-summary(hyp2F, test = Ftest())
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html



## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```

Change in G1 from T2 to T3: `r round(or[3,3],3)`

Change in G2 from T2 to T3: `r round(or[6,3],3)`

p=`r round(s$test$pvalue,4)`. We have `r evidence` that the change from T1 to T3 is different for G1 and G2 


##### Change from T1-T3



```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, 0,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2, test = Ftest())
summary(hyp2F, test = Ftest())
s<-summary(hyp2F, test = Ftest())
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html


## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```


Change in G1 from T1 to T3: `r round(or[2,3],3)`

Change in G2 from T1 to T3: `r round(or[5,3],3)`


p=`r round(s$test$pvalue,4)`. We have `r evidence` that the change from T1 to T3 is different for G1 and G2 


#### Residuals

Assumptions of the model are met when analysing the log-trasformed data.

```{r, echo=FALSE, warning=FALSE}
plot(result3,form = resid(., type = "pearson") ~ predict(., re.form = NA) )


qqnorm(residuals(result3))
qqline(residuals(result3), col="red")

# plot residuals against predictors
dat <- na.omit(dat)
ggplot(data.frame(x1=dat$group,pearson=residuals(result3,type="pearson")),
       aes(x=x1,y=pearson)) +
  geom_point() +
  theme_bw()

ggplot(data.frame(x2=dat$time,pearson=residuals(result3,type="pearson")),
       aes(x=x2,y=pearson)) +
  geom_point() +
  theme_bw()


```

```{r}
rm(list = ls())
```
