---
title: "PCIT-CoS-Waitlist RCT Paper 1: SSP Insecurity"
author: "Nancy Briggs"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 6
    toc_float: true
---


### SSP Insecurity {.tabset .tabset-fade .tabset-pills}

Output run on `r Sys.Date()`


variables are SSP_Pre_three_way_secure_or_insecure & SSP_Post_three_way_secure_or_insecure

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=1)
library(lme4)

library(tidyverse)
library(tidyselect)
library(SASmarkdown)
library(nlme)
library(emmeans)
library(multcomp)
library(kableExtra)
library(ggplot2)
library(lmerTest)
library(effectsize)
library(boot)

options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)

```


```{r, echo=FALSE}
# Data 
# Import data
# recoding for CBCL and totals done in SPSS before bringing data in to R

data <- read.csv("../data_prep.csv", fileEncoding="UTF-8-BOM")
names(data) <- tolower(names(data))

# keep relevant demographic and outcome variables

data <- data %>%
  dplyr::select(id:treatment_time_mins, dobchild, sexchild, numbhousehold, incfamily,
                #DPICS Labelled Praise
                parent_lp_perc_total_time_1, parent_lp_perc_total_time_2, parent_lp_perc_total_time_3,
                #DPICS Negative Talk
                parent_nta_perc_total_time_1, parent_nta_perc_total_time_2, parent_nta_perc_total_time_3,
                #BCAP
                prebcap_risk, post_bcap_risk, fu_bcap_risk,
                #BCAP Lie scale
                prebcap_lie, post_bcap_lie, fu_bcap_lie,
                
                # Child Ext Behav
                cbclext_pre_total,cbclext_post_total,cbclext_fu_total,

                # CBCL Internalising Behav
                cbclint_pre_total,cbclint_post_total,cbclint_fu_total,
                
                #Child Compliance
                total_co_perc_time_1,total_co_perc_time_2,total_co_perc_time_3,
                
                # others, but not essential for paper
                #3. Child Temperament (ECBQ)
                #4. Child social-emotional functioning - 
                #5. Child social-emotional functioning - #2: Child Devereux Early Childhood Assessment (DECA)
                
                #CQQ self Effic.
                pre_ccq_a_topse_mean, post_ccq_a_topse_mean ,fu_ccq_a_topse_mean, 

                # CCQ Caregiving Helplessness
                pre_chqd_sum, post_chqd_sum, fu_chqd_sum,
                
                # CCQ Hostil Parenting
                pre_ccqc_sum, post_ccqc_sum, fu_ccqc_sum,
                
                #1: Brief Infant Toddler Social Emotional Assessment (BITSEA)
                # Problem
                pre_bitseaprob_sum, post_bitseaprob_sum, fu_bitseaprob_sum,
                # Competence
                pre_bitseacomp_sum, post_bitseacomp_sum,fu_bitseacomp_sum,

                #SSP attachemnt Insecurity
                ssp_pre_three_way_secure_or_insecure , 
                ssp_post_three_way_secure_or_insecure
                )

data <- data %>%
  mutate(ssp_pre_three_way_secure_or_insecure=ifelse(ssp_pre_three_way_secure_or_insecure==999,NA,ssp_pre_three_way_secure_or_insecure)) %>%
  mutate(ssp_post_three_way_secure_or_insecure=ifelse(ssp_post_three_way_secure_or_insecure==999,NA,ssp_post_three_way_secure_or_insecure))


```




#### Descriptives

```{r, echo=FALSE, warning=F}
#   ssp_pre_three_way_secure_or_insecure ,                 ssp_post_three_way_secure_or_insecure

desctable <- data %>%
  group_by(group,ssp_pre_three_way_secure_or_insecure) %>%
  summarise(n1 = sum(!is.na(ssp_pre_three_way_secure_or_insecure)) ) %>%
  filter(!is.na(ssp_pre_three_way_secure_or_insecure)) %>%
  pivot_wider(id_cols=group, values_from=n1, 
              names_from = ssp_pre_three_way_secure_or_insecure) %>%
  ungroup() %>%
  mutate(total=`0`+`1`) %>%
  mutate(pct0=`0`/total) %>%
  mutate(pct1=`1`/total) %>%
  mutate(time="Pre") %>%
  relocate(time, .after=group) %>%
  relocate(pct0, .after=`0`) %>%
  relocate(pct1, .after=`1`) 


desctable1 <- data %>%
  group_by(group,ssp_post_three_way_secure_or_insecure) %>%
  summarise(n1 = sum(!is.na(ssp_post_three_way_secure_or_insecure)) ) %>%
  filter(!is.na(ssp_post_three_way_secure_or_insecure)) %>%
  pivot_wider(id_cols=group, values_from=n1, 
              names_from = ssp_post_three_way_secure_or_insecure) %>%
  ungroup() %>%
  mutate(total=`0`+`1`) %>%
  mutate(pct0=`0`/total) %>%
  mutate(pct1=`1`/total) %>%
  mutate(time="Post") %>%
  relocate(time, .after=group) %>%
  relocate(pct0, .after=`0`) %>%
  relocate(pct1, .after=`1`) 

desctable <- desctable %>%
  bind_rows(desctable1)

write.csv(desctable,"desctable.csv", row.names = TRUE)

```



```{r, echo=FALSE}
desctable %>%
           kbl(digits=3) %>%
           kable_minimal()
```


#### Model & Results

A generalized linear mixed model (GLMM) with random effect for individual and fixed effects of group, time and the interaction was run. A binomial distribution with logit link was specified. Odds ratios were obtained for group differences and change within group.

```{r, echo=FALSE, warning=FALSE}
#  analysis
dat <- data %>%
  dplyr::select(id,group,ssp_pre_three_way_secure_or_insecure,
                ssp_post_three_way_secure_or_insecure) %>%
  pivot_longer( cols = c("ssp_pre_three_way_secure_or_insecure", "ssp_post_three_way_secure_or_insecure"),  
                names_prefix = "bitseaprob_", 
                names_to="t", values_to="outcome") %>%
  mutate(time = ifelse(t == "ssp_pre_three_way_secure_or_insecure",  
                       1,2)) %>%
  # replace with NA the one group 3 with a post value
  mutate(outcome=ifelse(group==3 & time==2,NA,outcome)) %>%
  filter(!is.na(outcome))



dat$time<- as.factor(dat$time)
dat$group<- as.factor(dat$group)

result0<-glmer(outcome ~ (1 | id), family=binomial, dat)
result1<-glmer(outcome ~ group + (1 | id), family=binomial, dat)
result2<-glmer(outcome ~ group + time + (1 | id), family=binomial, dat)
result3<-glmer(outcome ~ group*time + (1 | id), family=binomial, dat)

```





```{r, echo=FALSE}
anovatab<-anova(result0,result1,result2,result3)

anovatab <-anovatab %>%
           mutate(Effect = ifelse(npar == 3, "Intercept",
                           ifelse(npar == 5, "Group",
                           ifelse(npar == 7, "Time",
                           ifelse(npar == 10, "Group*Time",""))))) %>%
           relocate(Effect) 
anovatab %>%
           kbl(digits=3) %>%
           kable_styling()


## Extract pvaue for evidence / no evidence
evidence = ifelse(anovatab[4,9]<0.05, "evidence", "no evidence")
```



This is the ANOVA table for the model. p-values are likelihood ratio tests for adding the new term to the model just before. A p-value <0.05 indicates that adding that term results in better model fit. The most important p-value is the interaction p-value. Here, there is `r evidence` that the pattern of change over time is different among the groups (p=`r round(anovatab[4,9],3)`).




```{r}
summary(result3)
```



#### Estimated Means & Change

Estimated means for each Treatment and Time are presented. In addition, within-group
change between time points is presented with 95% CIs.


```{r child='bootstrapped_es_CI.Rmd'}

```

<!-- ```{r child='TestingHypotheses.Rmd'}
``` -->

#### Hypothesis 2

##### Difference between G1 and G2, change from T1 to T3


```{r, echo=FALSE}
result.change

      
confint(result.change.d)
or<-data.frame(result.change.d)
```




```{r, echo=F, include=T, eval=T}
## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,  0,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2)
summary(hyp2F)
s<-summary(hyp2F, test = Ftest())
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html



## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```

Change in G1 from T1 to T3: `r round(or[1,3],3)`

Change in G2 from T1 to T3: `r round(or[2,3],3)`

p=`r round(s$test$pvalue,4)`. We have `r evidence` that the change from T1 to T3 is different for G1 and G2 




#### Residuals

Assumptions of the model are met.

```{r, echo=FALSE, warning=FALSE}
library(DHARMa)
sim= simulateResiduals(result3)
resid = residuals(sim, quantileFunction = qnorm, outlierValues = c(-7,7))
plot(resid~predict(result3, re.form = NA,type="response"))


qqnorm(resid)
qqline(resid, col="red")

# plot residuals against predictors
dat <- na.omit(dat)
ggplot(data.frame(x1=dat$group,
                  pearson=resid),
       aes(x=x1,y=pearson)) +
  geom_point() +
  theme_bw()

ggplot(data.frame(x2=dat$time,
                  pearson=resid),
       aes(x=x2,y=pearson)) +
  geom_point() +
  theme_bw()


```

```{r}
rm(list = ls())
```