---
title: "Hypothesis Tests"
author: "Nancy Briggs"
date: "`r Sys.Date()`"
output: html_document
---


#### Hypothesis 1

We hypothesize that from baseline (‘T1’) to post-treatment/waitlist (‘T2’) there will be a gradient effect, whereby parent-child dyads who receive PCIT-T
will show the greatest gains and parent-child dyads who
receive no treatment will show the least gains. For the outcome, then, the 
null hypothesis is:


$H_0:  \text {mean change}(\text{PCIT-T}) = \text {mean change}(\text{CoS-P}) = \text {mean change}(\text{Waitlist})$
$H_A: \text {mean change}(\text{PCIT-T}) > \text {mean change}(\text{CoS-P}) > \text {mean change}(\text{Waitlist})$

This is a joint hypothesis test.

First, confirming that the ordering of the means is as hypothesized:

```{r, echo=FALSE}
result.change.d
```

So, the mean change for PCITT > COSP > Waitlist.

Now, to get the p-value associated with the joint hypothesis.


```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp1   <- matrix(c(0, 0,0, 0,0, -1,0,0 ),1)
# Hyp2 is 1 df test, 
hyp2   <- matrix(c(0, 0,0, 0,1, 0,1,0 ),1)

# hypF is combined 2df test. Yes!
h<-rbind(hyp1,hyp2)
hypF <- as.matrix(h)
GlobalF <- glht(result3, linfct = hypF, test = Ftest())
GlobalF$linfct
summary(GlobalF, test = Ftest())
s<-summary(GlobalF, test = Ftest())

#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html

## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```

p=`r round(s$test$pvalue,4)`. We have `r evidence` that this the change from T1 to T2 is greater for G1 then G2 and the change in G2 is greater than in G3.




#### Hypothesis 2

##### Change from T2-T3

There will be a gradient of size of change from T2 to T3 such that the change in the PCIT-T group > the change in the CoS-P group.  This isn't really a gradient; it's just 1 hypothesis.

First, the change in PCITT and COSP from time 2 to time 3:

```{r, echo=FALSE}
result.change

      
confint(result.change.d)
or<-data.frame(result.change.d)
```




```{r}
## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, -1,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2)
summary(hyp2F)
s<-summary(hyp2F, test = Ftest())
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html



## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```

Change in G1 from T2 to T3: `r round(or[3,3],3)`

Change in G2 from T2 to T3: `r round(or[6,3],3)`

p=`r round(s$test$pvalue,4)`. We have `r evidence` that the change from T1 to T3 is different for G1 and G2 


##### Change from T1-T3



```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, 0,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2, test = Ftest())
summary(hyp2F, test = Ftest())
s<-summary(hyp2F, test = Ftest())
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html


## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```


Change in G1 from T1 to T3: `r round(or[2,3],3)`

Change in G2 from T1 to T3: `r round(or[5,3],3)`


p=`r round(s$test$pvalue,4)`. We have `r evidence` that the change from T1 to T3 is different for G1 and G2 


