---
title: "Imputation"
author: "Nancy Briggs"
date: "`r Sys.Date()`"
params:
  response : "ptsd"
---

Imputation for 

1.	Sensitivity Parenting sensitivity 	NICHD-
Sensitivity

2.	Positive parenting verbalizations Positive parenting verbalizations	DPICS – LP 
3.	Negative parenting verbalizations Negative parenting verbalizations 	DPICS – NTA
4.	Parental reflectiveness Parental reflectiveness	DRFQ  CCQE
5.	BITSEA-competence
6.	CBCL-externalizing
7.	BITSEA-problem

```{r setup, include=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE , warnings=FALSE)
options(warnPartialMatchAttr = FALSE, 
  warnPartialMatchDollar = FALSE,
  warnPartialMatchArgs = FALSE)
```

```{r, message=FALSE, echo=F, warnings=F}
library(lme4)
library(mice)
library(tidyverse)
library(tidyselect)
library(emmeans)
library(kableExtra)
library(ggplot2)
library(lmerTest)
library(effectsize)
library(boot)
library(VIM)
library(emmeans)
library(broom.mixed)
library(huxtable)

```


```{r, eval=F, warning=F , message=FALSE}

library(lme4)
library(mice)
library(tidyverse)
library(tidyselect)
library(emmeans)
library(kableExtra)
library(ggplot2)
library(lmerTest)
library(effectsize)
library(boot)
library(VIM)
library(emmeans)
library(broom.mixed)
library(huxtable)

```



```{r, echo=FALSE}
# Data 
# Import data
# recoding for CBCL and totals done in SPSS before bringing data in to R

data <- read.csv("../data_prep.csv", fileEncoding="UTF-8-BOM")
names(data) <- tolower(names(data))


# keep relevant demographic and outcome variables
data <- data %>%
  dplyr::select(id, agechild, sexchild, numbhousehold, incfamily,group,
        agemo, edumo, ethnicitymo,
        agefa, edufa, ethnicityfa, marpa, incfamily,
                      # Maternal reflective functioning 
                pre_ccqe_sum, post_ccqe_sum, fu_ccqe_sum,          
                
                                # Child Ext Behav
                cbclext_pre_total,cbclext_post_total,cbclext_fu_total,
                                #Observed maternal sensitivity 
                nichd_total_sensitivity, post_nichd_total_sensitivity, fu_nichd_total_sensitivity,
                    # Problem
                pre_bitseaprob_sum, post_bitseaprob_sum, fu_bitseaprob_sum,
                # Competence
                pre_bitseacomp_sum, post_bitseacomp_sum,fu_bitseacomp_sum
                )  %>%
  # set some 999 & 888 values to NA
  mutate_at(vars(sexchild, numbhousehold, incfamily, edumo, edufa, ethnicitymo, ethnicityfa, marpa), 
            .funs = function(x) replace(x, which(x == 999 | x == 888), NA)) 


### import total time
time <- read.csv("../data_prep_DPiCStime.csv", fileEncoding="UTF-8-BOM")
names(time) <- tolower(names(time))
time <- time %>%
  #DPICS Labelled Praise DPICS Negative Talk
  dplyr::select(id, parent_total_verbalisations_time_1, parent_total_lp_time_1, 
                  parent_total_nta_time_1,
                  parent_total_verbalisations_time_2,  parent_total_lp_time_2, 
                  parent_total_nta_time_2,
                  parent_total_verbalisations_time_3,parent_total_lp_time_3, 
                  parent_total_nta_time_3)
# total verbalisations goes with lp & nta in GLMM



# Join
data <- left_join(data, time, by="id")


# nichd_total_sensitivity, post_nichd_total_sensitivity, fu_nichd_total_sensitivity
# One person on Waitlist has a value for FU.  set to NA
data$fu_nichd_total_sensitivity[data$group==3] <- NA
```


```{r, eval=F, echo=F, include=F}



```

```{r}
mi<-data %>%
  relocate(group, .after="id") %>%
  relocate(parent_total_verbalisations_time_2, .after="parent_total_verbalisations_time_1") %>%
  relocate(parent_total_verbalisations_time_3, .after="parent_total_verbalisations_time_2") %>%
  relocate(parent_total_nta_time_2, .after="parent_total_nta_time_1") %>%
  relocate(parent_total_nta_time_3, .after="parent_total_nta_time_2") %>%
  relocate(parent_total_lp_time_2, .after="parent_total_lp_time_1") %>%
  relocate(parent_total_lp_time_3, .after="parent_total_lp_time_2") 
```



# Multiple Imputation


```{r, echo=F, include=F, eval=T}

# create 
#mi <- mi %>% 
#  mutate(time_group_int = interaction(rc_trial_time,group_type_id)) 

## now we impute
# initialise
ini <- mice(mi, maxit = 0)
# check for problems with variables, e.g. multicoliniarity etc
ini$loggedEvents

# remove problem variables
# mi_forimp <- mi %>% 
#   select(-ini$loggedEvents$out) 
mi_forimp <- mi

# re - initialize and check
# ini <- mice(mi_forimp, maxit = 0)
# ini$loggedEvents
 
```

## Checking correlations

Look at correlations among the variables. 

```{r}
# check correlations
mcl_dat = mi %>%
  dplyr::select(-id)
Xmat <- model.matrix(~., data = mcl_dat )
corrplot::corrplot(cor(Xmat[,-1]))

```


## Checking missing data patterns

```{r, echo=F}

# Plotting the missing data patterns

temp1<- data %>%
  dplyr::select(-id)
aggr =   aggr(temp1, col=mdc(1:2), numbers=TRUE, sortVars=TRUE, labels=names(temp1), cex.axis=.7, gap=3, ylab=c("Proportion of missingness","Missingness Pattern"))
aggr


```



```{r, echo=F}




# make id cluster variable
pred <- ini$pred
pred[-which(colnames(pred) == "id"), "id"] = 0  #check this is correct each time
#pred[-which(colnames(pred) == "time"), "time"] = 0  #we have an interaction that covers these
#pred[-which(colnames(pred) == "group"), "group"] = 0  #we have an interaction that covers these
# Make demographic vars predictors of all others, and no outcomes as predictors
pred[,c(14:37)]<-0



```





## Imputation details

Multiple imputation by chained equations was performed using the `mice` package in R. For the primary outcome DERS, there was 11% missing data at post and 16% at follow up.

The imputation model included as predictors: group, time, group*time, age and sex.  Missing values for the following outcomes were imputed:

  - DERS
  - Pain
  - PROMIS T
  - BDI
  - STAI
  - Well Being
  - PTSD
  - Stress T
  - Sleep Index 2


Twenty-five datasets were imputed, using the "2l.pan" method in `mice` to account for repeated measures within person.

The unadjusted mixed model was run on each imputed dataset, and all results, including the group difference in change (primary outcome) and effect size were aggregated using Rubin's Rules (1987).


```{r, eval=F}

# Create custom imputation for waitlist==3 & time 3 measures
# make a logical vector to say where you do and
# don't want imputing to occur
here <- mi_forimp %>% 
  mutate(across(c(fu_ccqe_sum,cbclext_fu_total,fu_nichd_total_sensitivity,fu_bitseaprob_sum,fu_bitseacomp_sum,
                  parent_total_verbalisations_time_3,parent_total_lp_time_3,parent_total_nta_time_3), 
                ~ case_when(group==3  ~ FALSE,
                            group!=3  ~ TRUE))) %>%
  mutate(across(c(pre_ccqe_sum,cbclext_pre_total,nichd_total_sensitivity,pre_bitseaprob_sum,pre_bitseacomp_sum,
                  parent_total_verbalisations_time_1,parent_total_lp_time_1,parent_total_nta_time_1),
                ~ case_when(group<=3  ~ TRUE))) %>%
  mutate(across(c(post_ccqe_sum,cbclext_post_total,post_nichd_total_sensitivity,post_bitseaprob_sum,post_bitseacomp_sum,
                  parent_total_verbalisations_time_2,parent_total_lp_time_2,parent_total_nta_time_2) ,
                ~ case_when(group<=3  ~ TRUE))) %>%
  mutate(across(c(agechild,sexchild, numbhousehold,incfamily, agemo, agefa, edumo, edufa, ethnicitymo, ethnicityfa, marpa) ,
                ~ case_when(group<=3  ~ TRUE))) %>%
  mutate(group = FALSE)


meth <- ini$meth
# normal for most outcomes
meth[c(14:28)] <- "norm"
# PMM for NTA and LP
meth[c(29:37)] <- "pmm"
# pmm
meth[c(3:13)] <- "pmm"
 
# impute
imp <- mice(mi_forimp, m = 25,  where = here, meth=meth, pred = pred, visit = "monotone",
            print = TRUE, maxit = 20)
save(imp, file ="imputed_full_allvars_long_pan") #
```




### Evaluation of imputations

```{r}
 
 # load previously imputed data
load(file ="imputed_full_allvars_long_pan")

# check for problems
imp$loggedEvents

# plot convergence diagnostic
plot(imp)
```



There were no problems with the imputations, and all imputations converged.


```{r}
# density plots
 densityplot(imp)
```



Imputed distributions look similar to the observed distributions.  This is good.



## Imputed Estimate of group difference in change in DERS to Follow up


Summary of two of the imputed datasets

```{r, echo=F}
load( file ="imputed_full_allvars_long_pan") # loads imp_mids

imputed_data<-complete(imp, action = "long", include = TRUE)
group <- imputed_data %>%
  filter(.imp==0) %>%
  dplyr::select(.id:group)
imputed_data<-imputed_data %>%
  dplyr::select(!c(id, group)) %>%
  left_join(group, by=c(".id")) %>%
  relocate(id, .after=".id") %>%
  relocate(group, .after="id")


# Make long
long1 <- imputed_data %>%
  dplyr::select(.imp,.id, id, pre_ccqe_sum, post_ccqe_sum, fu_ccqe_sum) %>%
  pivot_longer(cols=c(pre_ccqe_sum, post_ccqe_sum, fu_ccqe_sum),
                names_to="t", values_to="ccqe")%>%
  mutate(time = ifelse(t == "pre_ccqe_sum",  1,
                ifelse(t == "post_ccqe_sum", 2, 3))) %>%
  dplyr::select(!t)
long2 <- imputed_data %>%
  dplyr::select(.imp,.id, id, nichd_total_sensitivity, post_nichd_total_sensitivity, fu_nichd_total_sensitivity) %>%
  pivot_longer(cols=c( nichd_total_sensitivity, post_nichd_total_sensitivity, fu_nichd_total_sensitivity),
                names_to="t", values_to="sensitivity")%>%
  mutate(time = ifelse(t == "nichd_total_sensitivity",  1,
                ifelse(t == "post_nichd_total_sensitivity", 2, 3))) %>%
  dplyr::select(!t) %>%
  full_join(long1, by=c(".imp", ".id", "id","time") )
long1 <- imputed_data %>%
  dplyr::select(.imp,.id, id,  pre_bitseaprob_sum, post_bitseaprob_sum, fu_bitseaprob_sum) %>%
  pivot_longer(cols=c( pre_bitseaprob_sum, post_bitseaprob_sum, fu_bitseaprob_sum),
                names_to="t", values_to="bitseaprob")%>%
  mutate(time = ifelse(t == "pre_bitseaprob_sum",  1,
                ifelse(t == "post_bitseaprob_sum", 2, 3))) %>%
  dplyr::select(!t) %>%
  full_join(long2, by=c(".imp", ".id", "id","time") )
long2 <- imputed_data %>%
  dplyr::select(.imp,.id, id, pre_bitseacomp_sum, post_bitseacomp_sum,fu_bitseacomp_sum) %>%
  pivot_longer(cols=c( pre_bitseacomp_sum, post_bitseacomp_sum,fu_bitseacomp_sum),
                names_to="t", values_to="bitseacomp")%>%
  mutate(time = ifelse(t == "pre_bitseacomp_sum",  1,
                ifelse(t == "post_bitseacomp_sum", 2, 3))) %>%
  dplyr::select(!t) %>%
  full_join(long1, by=c(".imp", ".id", "id","time") )
long1 <- imputed_data %>%
  dplyr::select(.imp,.id, id,  parent_total_nta_time_1, parent_total_nta_time_2, parent_total_nta_time_3) %>%
  pivot_longer(cols=c( parent_total_nta_time_1, parent_total_nta_time_2, parent_total_nta_time_3),
                names_to="t", values_to="parent_nta")%>%
  mutate(time = ifelse(t == "parent_total_nta_time_1",  1,
                ifelse(t == "parent_total_nta_time_2", 2, 3))) %>%
  dplyr::select(!t) %>%
  full_join(long2,  by=c(".imp", ".id", "id","time") )
long2 <- imputed_data %>%
  dplyr::select(.imp,.id, id,   parent_total_lp_time_1, parent_total_lp_time_2, parent_total_lp_time_3) %>%
  pivot_longer(cols=c( parent_total_lp_time_1, parent_total_lp_time_2, parent_total_lp_time_3),
                names_to="t", values_to="parent_lp")%>%
  mutate(time = ifelse(t == "parent_total_lp_time_1",  1,
                ifelse(t == "parent_total_lp_time_2", 2, 3))) %>%
  dplyr::select(!t) %>%
  full_join(long1,  by=c(".imp", ".id", "id","time") )
long1 <- imputed_data %>%
  dplyr::select(.imp,.id, id,   parent_total_verbalisations_time_1, parent_total_verbalisations_time_2, parent_total_verbalisations_time_3) %>%
  pivot_longer(cols=c( parent_total_verbalisations_time_1, parent_total_verbalisations_time_2, parent_total_verbalisations_time_3),
                names_to="t", values_to="verbal")%>%
  mutate(time = ifelse(t == "parent_total_verbalisations_time_1",  1,
                ifelse(t == "parent_total_verbalisations_time_2", 2, 3))) %>%
  dplyr::select(!t) %>%
  full_join(long2, by=c(".imp", ".id", "id","time")  )

long <- imputed_data %>%
  dplyr::select(.imp,.id, id,  agechild, sexchild, numbhousehold, incfamily,group,
        agemo, edumo, ethnicitymo,
        agefa, edufa, ethnicityfa, marpa, incfamily,
                cbclext_pre_total,cbclext_post_total,cbclext_fu_total) %>%
  pivot_longer(cols=c(  cbclext_pre_total,cbclext_post_total,cbclext_fu_total),
                names_to="t", values_to="cbclext")%>%
  mutate(time = ifelse(t == "cbclext_pre_total",  1,
                ifelse(t == "cbclext_post_total", 2, 3))) %>%
  dplyr::select(!t) %>%
  full_join(long1,  by=c(".imp", ".id", "id","time") )  %>%
  #create  id variable in putting data back into mids object
  mutate(time_id = paste(time,.id,id, sep="_")) 


rm(long1, long2)



# now convert back to a mids object for modelling
imp_mids <- as.mids(long, .id="time_id")

# and save
save(imp_mids, file ="imputed_long") # 

# 
# long %>% 
#   filter(.imp %in% 1:2) %>% 
#   group_by(.imp, group, time ) %>% 
#    summarize_at(vars(cbclext, ccqe), 
#                 .funs = list(Mean = mean, Sd = sd, N_obs = ~length(.)), na.rm=T)



```

