---
title: "Hypothesis Tests"
author: "Nancy Briggs"
date: "`r Sys.Date()`"
output: html_document
---

#### Hypothesis 1

We hypothesize that from baseline (‘T1’) to post-treatment/waitlist (‘T2’) there will be a gradient effect, whereby parent-child dyads who receive PCIT-T
will show the greatest gains and parent-child dyads who
receive no treatment will show the least gains. For the outcome, then, the 
null hypothesis is:

$H_0:  \text {mean change}(\text{PCIT-T}) = \text {mean change}(\text{CoS-P}) = \text {mean change}(\text{Waitlist})$
$H_A: \text {mean change}(\text{PCIT-T}) > \text {mean change}(\text{CoS-P}) > \text {mean change}(\text{Waitlist})$

This is a joint hypothesis test.

First, confirming that the ordering of the means is as hypothesized:

```{r, echo=FALSE}
result.change  <- emmeans(result3, "time", by = "group")
result.change.d <-pairs(result.change)
result.change.d
result.change.d1<-as.data.frame(result.change.d)
result.change.d1<- result.change.d1 %>%
  mutate(t1t2change=-1*estimate) %>%
  mutate(quality=ifelse(t1t2change<0, "decreased t1 to t2", 
                 ifelse(t1t2change>0, "increase t1 to t2", "did not change" )))

result.change.d1 %>%
  filter(contrast=="1 - 2") %>%
  dplyr::select(contrast, group, t1t2change, quality, p.value ) %>%
  kbl(digits=3) %>%
  kable_styling()


# ordering change.
## Extract pvalue for evidence / no evidence
orderincrease <- ifelse( result.change.d1$t1t2change[1]>result.change.d1$t1t2change[4] &
                         result.change.d1$t1t2change[4]>result.change.d1$t1t2change[7] , 
                 "PCITT > COS-P > Waitlist", "Order not satisfied" )
# orderdecrease <- ifelse( result.change.d1$estimate[1]<result.change.d1$estimate[4],
#          ifelse( result.change.d1$estimate[4]<result.change.d1$estimate[7] , 
#                  "PCITT > COS-P > Waitlist", "Order not satisfied" ))



```

Is the ordering of change satisfied? 

Greatest to least change from t1 to t2:  `r orderincrease`

Note that I interpreted an increase in scores as being "better".  If a decrease in scores is "better" then these change values should be interpreted in that light.


```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp1   <- matrix(c(0, 0,0, 0,0, 1,0,0 ),1)
# Hyp2 is 1 df test, 
hyp2   <- matrix(c(0, 0,0, 0,0, -1,1, 0 ),1)




# hypF is combined 2df test. Yes!
h<-rbind(hyp1,hyp2)
hypF <- as.matrix(h)
GlobalF <- glht(result3, linfct = hypF, test = Ftest())
GlobalF$linfct
summary(GlobalF, test = Ftest())
F<-summary(GlobalF, test = Ftest())

#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html


## Extract pvalue for evidence / no evidence
evidence = ifelse(F$test$pvalue<0.05, "evidence", "no evidence")

```


p= `r round(F$test$pvalue,3)`. We have `r evidence` that there is a difference in the change over from Baseline to Post amongst the groups.


#### Hypothesis 2

##### Change from T2-T3

There will be a gradient of size of change from T2 to T3 such that the change in the PCIT-T group > the change in the CoS-P group.  This isn't really a gradient; it's just 1 hypothesis.

First, the change in PCITT and COSP from time 2 to time 3:

```{r, echo=FALSE}
result.change

      
confint(result.change.d)
or<-data.frame(result.change.d)
```




```{r}
## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, -1,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2)
summary(hyp2F)
s<-summary(hyp2F, test = Ftest())
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html



## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```

Change in G1 from T2 to T3: `r round(or[3,3],3)`

Change in G2 from T2 to T3: `r round(or[6,3],3)`

p=`r round(s$test$pvalue,4)`. We have `r evidence` that the change from T1 to T3 is different for G1 and G2 


##### Change from T1-T3



```{r, echo=FALSE, warning=FALSE}

## Specify contrast matrices
# Hyp1 is 1 df test
# could use alternative="less" ir alternative="greater" in glht
hyp2   <- matrix(c(0, 0,0, 0,0, 0,0,1 ),1)
hyp2F <- glht(result3, linfct = hyp2, test = Ftest())
summary(hyp2F, test = Ftest())
s<-summary(hyp2F, test = Ftest())
#also see this webpage for more on contrasts https://rcompanion.org/rcompanion/h_01.html


## Extract pvaue for evidence / no evidence
evidence = ifelse(s$test$pvalue<0.05, "evidence", "no evidence")

```


Change in G1 from T1 to T3: `r round(or[2,3],3)`

Change in G2 from T1 to T3: `r round(or[5,3],3)`


p=`r round(s$test$pvalue,4)`. We have `r evidence` that the change from T1 to T3 is different for G1 and G2 

